### 2017-01-30 documents/ms.tex
initial

\documentclass[useAMS, usenatbib, preprint, 12pt]{aastex}
\usepackage{cite, natbib}
\usepackage{float}
\usepackage{epsfig}
\usepackage{cases}
\usepackage[section]{placeins}
\usepackage{graphicx, subfigure}
\usepackage{color}
\usepackage{bm}
\usepackage{enumerate}
\usepackage{pdflscape}

\newcommand{\columbia}{1}
\newcommand{\simons}{2}
\newcommand{\nyu}{3}
\newcommand{\cca}{4}
\newcommand{\mpia}{5}
\newcommand{\Kepler}{{\it Kepler}}
\newcommand{\kepler}{\Kepler}
\newcommand{\corot}{{\it CoRoT}}
\newcommand{\Ktwo}{{\it K2}}
\newcommand{\ktwo}{\Ktwo}
\newcommand{\TESS}{{\it TESS}}
\newcommand{\LSST}{{\it LSST}}
\newcommand{\Wfirst}{{\it Wfirst}}
\newcommand{\SDSS}{{\it SDSS}}
\newcommand{\PLATO}{{\it PLATO}}
\newcommand{\Teff}{$T_{\mathrm{eff}}$}
\newcommand{\teff}{$T_{\mathrm{eff}}$}
\newcommand{\FeH}{[Fe/H]}
\newcommand{\feh}{[Fe/H]}
\newcommand{\ie}{{\it i.e.}}
\newcommand{\eg}{{\it e.g.}}
\newcommand{\etal}{{\it et al.}}
\newcommand{\logg}{log \emph{g}}
\newcommand{\dnu}{$\Delta \nu$}
\newcommand{\numax}{$\nu_{\mathrm{max}}$}

\begin{document}

\title{A coherent and comprehensive dating system for stars in the Milky Way}

\author{%
    Ruth Angus\altaffilmark{\columbia,}\altaffilmark{\simons},
   David Hogg\altaffilmark{\nyu,}\altaffilmark{\mpia,}\altaffilmark{\cca},
    \etal
   % Melissa Ness\altaffilmark{\mpia},
   % David Kipping\altaffilmark{\columbia}, \etal
}

\altaffiltext{\columbia}{Department of Astronomy, Columbia
University, NY, NY}
\altaffiltext{\simons}{Simons Fellow, RuthAngus@gmail.com}
\altaffiltext{\nyu}{Center for Cosmology and Particle Physics, New York
University, NY, NY}
\altaffiltext{\mpia}{Max Planck Institute of Astronomy, Heidelberg, Germany}
\altaffiltext{\cca}{Center for Computational Astrophysics, Flatiron Institute,
NY, NY}

\begin{abstract}
Words go here.
\end{abstract}

\section{Introduction}
\label{sec:intro}

The processes behind the formation of the Milky Way and the objects within it
are some of the most elusive and complicated topics in astronomy today,
connected by a common theme: stellar ages.
Ages provide the key to understanding the evolution of all astrophysical
objects, but age, unfortunately, one of the most difficult stellar properties
to measure.
Different dating methods often produce inconsistent predictions for the age of
the same star.
This is in part because the underlying processes generating the evolution of
the observable properties are different and in part because our understanding
of the underlying physics is flawed or incomplete.
In table \ref{tab:dating_methods} we present a list of dating methods, the
main observables associated with that method, the underlying physics driving
the changing observables, the types of star the method applies to and the
empirical or physical nature of the model.

\begin{landscape}
\begin{table}
\begin{center}
\caption{A list of all dating methods with the types of stars they apply to
    and their effective or physical natures.}
\begin{tabular}{lcccc}
    Method & Observable & Underlying cause & Applicable to... & Effective or
    Physical model? \\
    \hline
    Rotation & $P_{\mathrm{rot}}$ & Magnetic activity & $<$ 4 Gyr, MS &
    Effective \\
    Activity & Radio, H$\alpha$ \& X-ray & Magnetic activity & $<$
    5 Gyr, low mass & Effective \\
    Time variability & Time variability & ? & ? & ? \\
    Isochrones & $T_{\mathrm{eff}}$, $M_v$ & Core/shell fusion &
    Subgiants/giants & Physical (model dependent) \\
    Asteroseismology & Frequencies & Core/shell fusion & Old, low
    mass, Giants & Physical (model dependent) \\
    White dwarf cooling & Cooling & $M_v$ & WDs (\& companions) &
    Physical (1 free parameter) \\
    $J_z$, $J_{\phi}$, $J_r$ & Position \& velocity & Dynamic heating & Disk
    stars & Physical (1 free parameter) \\
    Galactocentric position & Position & Formation history &
    All stars & Effective \\
    Coeval pairs & Position \& velocity & Disruption & $<$ 1 Gyr, disk &
    0 free parameters \\
    Open Clusters & Disruption & Positions \& velocities & $<$ 2 Gyr &
    0 free parameters \\
    Exoplanet dynamics & Stability \& eccentricity & Interactions & $<$ 500
    Myrs & 0 free parameters \\
    Universe age & All of the above & Finite age of the Universe &
    All stars & 0 free parameters \\
    C/N ratio & C/N ratio & Convective processes & Red giants, A \& F &
    Effective \\
    Chemical abundances & Abundances & ISM enrichment & $>$ 8 Gyr?
    & Effective \\
    Lithium abundance & L abundance & $L$ depletion & $<$ 500
    Myrs? & Effective \\
    Nuclear isotopes & ? & ? & ? & ? \\
    Unusual HRD objects & $T_{\mathrm{eff}}$, $M_v$ & Mergers, ? &
    Special cases & ? \\
\end{tabular}
\end{center}
\end{table}
\label{tab:dating_methods}
\end{landscape}

\subsection{Rotation-Dating}
\label{sec:rotation}

Main sequence (MS) stars comprise the majority of our galaxy but their ages
are notoriously difficult to measure.
Their positions on the HR diagram don't change significantly during their
    hydrogen burning lifetimes, a fact that is convenient for life on Earth
    but inconvenient for galactic archaeologists.
Now, due to the abundance of rotation periods for MS stars provided by Kepler
    and to-be provided by TESS, LSST and Wfirst, rotation-dating is the most
    readily available, precise method for inferring stellar ages.
Rotation-dating works well for young stars but a question mark still hangs
    over its accuracy for stars older than the Sun.
Recent results show that old \kepler\ asteroseismic stars rotate more rapidly
    than expected given their age \citep[\eg][]{Angus2015, Vansaders2016,
    Metcalfe2016}.
This has been attributed to an evolving magnetic dynamo: as stars reach an
    critical Rossby number (the ratio of rotation period to the convective
    overturn timescale), their magnetic field `switches off' and stars
    maintain a consistent rotation period after that time.
Whilst this physical explanation produces a model that fits the data, it
    is driven by observations, not theory, and other explanations could
    provide an answer.
The data sets typically used to test the age-rotation relations are highly
    heterogeneous and each set has its own detection and selection biases.
For example, asteroseismology favours quiet stars whereas rotation periods are
    easiest to measure for active stars.

% A history of rotation-dating.
The phenomenon of magnetic braking in MS stars was first observed almost fifty
years ago by \citet{Skumanich1972} who observed that the rotation periods of
the Sun and young cluster stars seemed to decay with the square-root of time.
Later, a mass-dependence was added to the relation between age and rotation
period --- less massive stars lose angular momentum faster than more massive
ones.
\citet{Kawaler1988} derived a formalism for this angular momentum loss and his
relation depended on the mass loss rate, the ....
More recently, \citet{Barnes2003} demonstrated that a simple relation could be
used to describe `gyrochronology', the method of rotation-dating, and further
works \citep[\eg][]{Barnes2007, Mamajek2008, Barnes2010, Meibom2011},
continue to demonstrate that the relation between rotation period and age
holds true while theorists \citep[\eg][]{Matt2012, Epstein2014} modify and
extend the efforts to produce physical models of this phenomenon.

\section{Method}

\section{Results and Discussion}

\section{Conclusions}

% acknowledgements
This research was funded by the Simons Foundation.
Some of the data presented in this paper were obtained from the Mikulski
Archive for Space Telescopes (MAST).
STScI is operated by the Association of Universities for Research in
Astronomy, Inc., under NASA contract NAS5-26555.
Support for MAST for non-HST data is provided by the NASA Office of Space
Science via grant NNX09AF08G and by other grants and contracts.
This paper includes data collected by the Kepler mission. Funding for the
Kepler mission is provided by the NASA Science Mission directorate.

\bibliographystyle{plainnat}
\bibliography{chronometer}

\end{document}

### 2017-02-06 documents/ms.tex
adding words on the subway

@@
 
+\newcommand{\racomment}[1]{{\color{red}#1}}
+
+% Affiliations
 \newcommand{\columbia}{1}
@@
 \newcommand{\mpia}{5}
+
+% Missions
 \newcommand{\Kepler}{{\it Kepler}}
@@
 \newcommand{\TESS}{{\it TESS}}
+\newcommand{\tess}{{\it TESS}}
 \newcommand{\LSST}{{\it LSST}}
+\newcommand{\lsst}{{\it LSST}}
 \newcommand{\Wfirst}{{\it Wfirst}}
+\newcommand{\wfirst}{{\it Wfirst}}
 \newcommand{\SDSS}{{\it SDSS}}
+\newcommand{\sdss}{{\it SDSS}}
 \newcommand{\PLATO}{{\it PLATO}}
+\newcommand{\plato}{{\it PLATO}}
+\newcommand{\TGAS}{{\it TGAS}}
+\newcommand{\tgas}{{\it TGAS}}
+\newcommand{\APOGEE}{{\it APOGEE}}
+\newcommand{\apogee}{{\it APOGEE}}
+\newcommand{\gaia}{{\it Gaia}}
+\newcommand{\Gaia}{{\it Gaia}}
+
+% Stellar parameters
 \newcommand{\Teff}{$T_{\mathrm{eff}}$}
@@
 \newcommand{\feh}{[Fe/H]}
+\newcommand{\logg}{log \emph{g}}
+\newcommand{\dnu}{$\Delta \nu$}
+\newcommand{\numax}{$\nu_{\mathrm{max}}$}
+
+% Common expressions
 \newcommand{\ie}{{\it i.e.}}
@@
 \newcommand{\etal}{{\it et al.}}
-\newcommand{\logg}{log \emph{g}}
-\newcommand{\dnu}{$\Delta \nu$}
-\newcommand{\numax}{$\nu_{\mathrm{max}}$}
+
+% Paper-specific
+\newcommand{\nkictags}{2000}
 
@@
 \begin{abstract}
-Words go here.
+% Aims
+Stellar age underpins a wide range of astronomical applications, yet
+remains one of the most challenging properties to infer.
+% Method
+Competing dating methods with different observational inputs often produce
+    wildly inconsistent age estimates.
+In this work we attempt to combine multiple methods into one coherent model
+    and calibrate that model using galactic dynamics --- a well understood
+    physical process.
+We identify a sample of stars with at least two of the following dating
+    methods available: rotation-dating, isochronal-dating,
+    asteroseismic-dating and dynamical-dating.
+% Results
 \end{abstract}
@@
 Ages provide the key to understanding the evolution of all astrophysical
-objects, but age, unfortunately, one of the most difficult stellar properties
-to measure.
+objects, but age is, unfortunately, one of the most difficult stellar
+properties to measure.
 Different dating methods often produce inconsistent predictions for the age of
@@
 of the underlying physics is flawed or incomplete.
-In table \ref{tab:dating_methods} we present a list of dating methods, the
-main observables associated with that method, the underlying physics driving
-the changing observables, the types of star the method applies to and the
-empirical or physical nature of the model.
+The various available dating methods can be categorised by the underlying
+physical process they trace --- the radial extent of the hydrogen-burning core
+or the evolving magnetic dynamo for example.
+In addition, they can be classified by their level of empiricism, \ie\ the
+number of free parameters that need to tuned when calibrating these models.
+The physics behind orbital heating, for example, is very well understood and
+has only one tunable parameter.
+On the other hand, magnetic activity evolution is poorly understood and the
+dating methods that are based on it are often entirely empirical.
+In table \ref{tab:dating_methods} we provide an overview of various dating
+methods, the main observables associated with them, the underlying physics
+driving the changing observables, the types of star the method applies to and
+the empirical or physical nature of the model.
 
@@
 
+In this paper we focus on four of these dating methods: rotation,
+asteroseismology, isochrones and galactic dynamics.
+A brief overview of each of these methods is provided in below.
+
 \subsection{Rotation-Dating}
@@
 
+Not only do a number of dating methods exists, several different models are
+often available for the same dating method.
+In the case of rotation-dating...
+
+\racomment{What happens when you can't measure a rotation period?}
+\racomment{What about the difference in physics that you are probing?}
+
+\subsection{Asteroseismology}
+
+\subsection{Isochrones}
+
+\subsection{Galactic Dynamics}
+
+This paper is laid out as follows.
+% In \textsection \ref{sec:dating_methods} we provide an overview of each dating
+% method used in this paper and demonstrate their inconsistency in age
+% predictions.
+In \textsection \ref{sec:data} we describe the data used to calibrate our new
+model, which is described in \textsection \ref{sec:method}.
+In \textsection \ref{sec:results} we present a new catalogue of ages.
+
+\section{Data}
+\label{sec:data}
+We assembled a sample of stars that have at least two of the four dating
+methods available.
+Our main asteroseismic sample targets comprise the $~$500 short cadence
+\kepler\ prime targets.
+These stars all have light curves and therefore rotation periods (or potential
+rotation periods).
+The majority are also in Gaia DR1, therefore have parallaxes, positions and
+2-dimensional velocities, which can provide constraints on their vertical
+actions.
+A subset of these stars have spectroscopic information and therefore make
+excellent isochronal targets.
+Our main rotation sample is the \nkictgas\ stars in both the \kepler\ input
+catalogue (KIC) and \TGAS.
+These stars have potential rotation periods, parallaxes, positions and
+2-dimensional velocities.
+Our main dynamics data set is the intersection of the \apogee\ and \tgas\
+catalogues.
+The majority of these stars have ages from {\tt the cannon} \cite{Ness2015},
+plus \gaia\ parallaxes, positions and proper motions as well as radial
+velocities, providing vertical actions calculated from the full 3-dimensional
+phase-space.
+They are mostly red giants.
+
 \section{Method}
+\label{sec:method}
+
+In order to establish the level of inconsistency between dating methods, we
+first compare predictions for each of the above data sets.
+This is purely an illustrative test as these dating methods are not always
+independent, for example many rotation-dating models are calibrated using
+asteroseismic age predictions.
+For the asteroseismic sample we find that
 
 \section{Results and Discussion}
+\label{sec:results}
 
@@
 Kepler mission is provided by the NASA Science Mission directorate.
+\racomment{Add TGAS \& Apogee acknowledgements.}
 

### 2017-02-06 documents/ms.tex
adding table of data sets

@@
 \newcommand{\kepler}{\Kepler}
+\newcommand{\kic}{\KIC}
+\newcommand{\KIC}{\KIC}
 \newcommand{\corot}{{\it CoRoT}}
@@
 \newcommand{\Gaia}{{\it Gaia}}
+\newcommand{\rave}{{\it RAVE}}
+\newcommand{\Rave}{{\it RAVE}}
 
@@
 % Paper-specific
-\newcommand{\nkictags}{2000}
+\newcommand{\nkictgas}{2000}
 
@@
     Ruth Angus\altaffilmark{\columbia,}\altaffilmark{\simons},
-   David Hogg\altaffilmark{\nyu,}\altaffilmark{\mpia,}\altaffilmark{\cca},
+   David Hogg\altaffilmark{\nyu,}\altaffilmark{\cca,}\altaffilmark{\mpia},
     \etal
@@
     Special cases & ? \\
+    Twins & ? & ? & ? & ? \\
 \end{tabular}
@@
 2-dimensional velocities.
-Our main dynamics data set is the intersection of the \apogee\ and \tgas\
-catalogues.
-The majority of these stars have ages from {\tt the cannon} \cite{Ness2015},
-plus \gaia\ parallaxes, positions and proper motions as well as radial
-velocities, providing vertical actions calculated from the full 3-dimensional
-phase-space.
-They are mostly red giants.
+Some of these stars, particularly those with confirmed planets, also have
+spectroscopic information and could provide isochronal ages.
+Our main isochronal data set is the intersection of the \apogee\ and \rave\
+catalogues with \tgas.
+Relatively precise isochronal ages can be inferred from the spectroscopic
+stellar properties of these stars, the majority which have ages from {\it the
+cannon} \citep{Ness2015, Casey2016}.
+They also have \gaia\ parallaxes, positions and proper motions as well as
+radial velocities, providing vertical actions calculated from the full
+3-dimensional phase-space.
+
+\begin{landscape}
+\begin{table}
+\begin{center}
+\caption{A summary of the data sets used in this project}
+\begin{tabular}{lccc}
+    Sample & Catalogue intersection &  Methods available \\
+    \hline
+    Asteroseismic & \citet{Chaplin2014} \& TGAS & Asteroseismology, rotation,
+    isochrones, dynamics \\
+    Rotational & KIC \& TGAS & Rotation, isochrones, dynamics \\
+    Isochronal & RAVE, APOGEE \& TGAS & Isochrones, dynamics \\
+\end{tabular}
+\end{center}
+\end{table}
+\label{tab:data}
+\end{landscape}
 

### 2017-02-10 chronometer/action_age_evolution.py
initial

@@ -1,0 +1,56 @@
+# # The Action-age relation
+
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+
+from galpy.potential import MWPotential2014
+from galpy.actionAngle import actionAngleStaeckel, actionAngleAdiabatic
+from galpy.df import quasiisothermaldf
+
+
+def calc_dispersion(time, sz0, t1, tm, beta, R0, Rc, Rsz):
+    return sz0 * ((time + t1)/(tm + t1))**beta * np.exp((R0 - Rc)/Rsz)
+
+
+if __name__ == "__main__":
+    time = np.linspace(0, 14, 100)
+    sz0 = 50.
+    sr0 = 50.
+    t1 = .1
+    tm = 10.
+    beta = .33
+    R0 = 1.
+    Rc = 1.
+    hsz = 9.
+    hsr = 9.
+
+    sigma_z = calc_dispersion(time, sz0, t1, tm, beta, R0, Rc, hsz)
+    sigma_r = calc_dispersion(time, sr0, t1, tm, beta, R0, Rc, hsr)
+
+    plt.clf()
+    plt.plot(time, sigma_z)
+    plt.savefig("test")
+
+    # Bovy example
+    # aA = actionAngleAdiabatic(pot=MWPotential2014, c=True)
+    # hr = 1/3.  # radial scale length
+    # sr = .2  # radial velocity dispersion at the solar radius (time dependent)
+    # sz = .1  # vertical velocity dispersion at the solar radius (time dependent)
+    # hsr = 1.  # radial-velocity-dispersion scale length
+    # hsz = 1.  # vertial-velocity-dispersion scale length
+
+    aA = actionAngleStaeckel(pot=MWPotential2014, delta=0.45, c=True)
+    solar_radius = 8.
+    hr = 2.68/solar_radius
+    # sr = 34.
+    # sz = 25.1
+
+    meanjzs = np.ones(len(sigma_z))
+    for i, sz in enumerate(sigma_z):
+        qdf = quasiisothermaldf(hr, sigma_r[i], sz, hsr, hsz,
+                                pot=MWPotential2014, aA=aA, cutcounter=True)
+        meanjzs.append(qdf.meanjz(1., .125))
+    print(qdf.estimate_hr(1.))
+    print(qdf.meanvz(1., 0.))
+    print(qdf.meanjz(1., .125))

### 2017-02-10 chronometer/action_age_evolution.py
words on the subway

@@ -7,6 +7,7 @@ import pandas as pd
 from galpy.potential import MWPotential2014
 from galpy.actionAngle import actionAngleStaeckel, actionAngleAdiabatic
 from galpy.df import quasiisothermaldf
+aA = actionAngleStaeckel(pot=MWPotential2014, delta=0.45, c=True)
 
 
 def calc_dispersion(time, sz0, t1, tm, beta, R0, Rc, Rsz):
@@ -24,6 +25,12 @@ if __name__ == "__main__":
     Rc = 1.
     hsz = 9.
     hsr = 9.
+    solar_radius = 8.
+    hr = 2.68/solar_radius
+
+    # Today
+    sr = 34.
+    sz = 25.1
 
     sigma_z = calc_dispersion(time, sz0, t1, tm, beta, R0, Rc, hsz)
     sigma_r = calc_dispersion(time, sr0, t1, tm, beta, R0, Rc, hsr)
@@ -33,24 +40,31 @@ if __name__ == "__main__":
     plt.savefig("test")
 
     # Bovy example
-    # aA = actionAngleAdiabatic(pot=MWPotential2014, c=True)
-    # hr = 1/3.  # radial scale length
-    # sr = .2  # radial velocity dispersion at the solar radius (time dependent)
-    # sz = .1  # vertical velocity dispersion at the solar radius (time dependent)
-    # hsr = 1.  # radial-velocity-dispersion scale length
-    # hsz = 1.  # vertial-velocity-dispersion scale length
+    aA = actionAngleAdiabatic(pot=MWPotential2014, c=True)
+    hrb = 1/3.  # radial scale length
+    srb = .2  # radial velocity dispersion at the solar radius (time dependent)
+    szb = .1  # vertical velocity dispersion at the solar radius (time dependent)
+    hsrb = 1.  # radial-velocity-dispersion scale length
+    hszb = 1.  # vertial-velocity-dispersion scale length
 
-    aA = actionAngleStaeckel(pot=MWPotential2014, delta=0.45, c=True)
-    solar_radius = 8.
-    hr = 2.68/solar_radius
-    # sr = 34.
-    # sz = 25.1
+    print(sr, srb)
+    qdf = quasiisothermaldf(hr, srb, szb, hsrb, hszb, pot=MWPotential2014,
+                            aA=aA, cutcounter=True)
+    print(qdf.meanjz(1., .125))
+    assert 0
 
[…]

### 2017-02-12 chronometer/staying_alive.py
adding staying_alive.py

@@ -1,0 +1,64 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.mlab as mlab
+
+
+def q(x, y):
+    g1 = mlab.bivariate_normal(x, y, 1.0, 1.0, -1, -1, -0.8)
+    g2 = mlab.bivariate_normal(x, y, 1.5, 0.8, 1, 2, 0.6)
+    return 0.6*g1+28.4*g2/(0.6+28.4)
+
+
+def metropolis_hastings(N, r, s):
+    """
+    params:
+    -------
+    N: (int)
+        Number of samples.
+    r: (tuple or list or array)
+        The parameters/position in probability space.
+    s: (int)
+        Thinning index
+
+    """
+
+    p = q(r[0], r[1])  # Initial probability.
+    samples = []
+    for i in np.arange(N):
+        rn = r + np.random.normal(size=2)  # Take a step in parameter space.
+        pn = q(rn[0], rn[1])  # calculate the probability.
+        if pn >= p:  # If the new probability is greater than the previous p
+            p = pn  # The probability is now the new one.
+            r = rn  # adopt the new parameters.
+        else:
+            u = np.random.rand()  # Otherwise, random number between 0 and 1.
+            if u < pn/p:  # if less than the ratio of new p to old p:
+                p = pn  # prob is now new prob.
+                r = rn  # adopt the new parameters.
+        if i % s == 0:  # If this is the 10*th step, append sample (thinning)
+            samples.append(r)
+    return np.array(samples)
+
+
+
+if __name__ == "__main__":
+
+    '''Metropolis Hastings'''
+    N = 100000  # N samples
+    s = 10  #
+    r = np.zeros(2)  # initialisation
+
+    samples = metropolis_hastings(N, r, s)
+
+    plt.clf()
+    plt.scatter(samples[:, 0], samples[:, 1], alpha=0.5, s=1)
+
+    '''Plot target distribution'''
+    dx = 0.01
+    x = np.arange(np.min(samples), np.max(samples), dx)
+    y = np.arange(np.min(samples), np.max(samples), dx)
+    X, Y = np.meshgrid(x, y)
+    Z = q(X, Y)
+    CS = plt.contour(X, Y, Z, 10)
+    plt.clabel(CS, inline=1, fontsize=10)
+    plt.savefig("test")

### 2017-02-12 chronometer/staying_alive.py
fitting a straight line

@@ -1,7 +1,53 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.mlab as mlab
+import scipy.misc as spm
+import corner
 
+
+#!/usr/bin/env python
+# Author:      Flávio Codeço Coelho
+# License:     GPL
+
+from math import *
+import numpy as np
+from matplotlib.pylab import *
+
+# def sdnorm(z):
+#     """
+#     Standard normal pdf (Probability Density Function)
+#     """
+#     return exp(-z*z/2.)/sqrt(2*pi)
+
+# n = 10000
+# alpha = 1
+# x = 0.
+# vec = []
+# vec.append(x)
+# innov = np.random.uniform(-alpha,alpha,n) #random inovation, uniform proposal distribution
+# for i in range(n):
+#     can = x + innov[i] #candidate
+#     aprob = min([1.,sdnorm(can)/sdnorm(x)]) #acceptance probability
+#     u = uniform(0,1)
+#     if u < aprob:
+#         x = can
+#         vec.append(x)
+
+# #plotting the results:
+# #theoretical curve
+# x = arange(-3,3,.1)
+# y = sdnorm(x)
+# subplot(211)
+# title('Metropolis-Hastings')
+# plot(vec)
+# subplot(212)
+
+# hist(vec, bins=30,normed=1)
+# plot(x,y,'ro')
+# ylabel('Frequency')
+# xlabel('x')
+# legend(('PDF','Samples'))
+# show()
 
 def q(x, y):
     g1 = mlab.bivariate_normal(x, y, 1.0, 1.0, -1, -1, -0.8)
@@ -9,24 +55,33 @@ def q(x, y):
     return 0.6*g1+28.4*g2/(0.6+28.4)
 
 
-def metropolis_hastings(N, r, s):
+def model(par, x):
+    return par[0] + par[1]*x
+
+
+def lnlike(par, x, y_obs, y_err):
+    y_mod = model(par, x)
+    return sum(-.5*((y_mod - y_obs)/y_err)**2)
+
+
+def metropolis_hastings(x, y, yerr, r, N, s, t):
     """
     params:
     -------
-    N: (int)
-        Number of samples.
     r: (tuple or list or array)
         The parameters/position in probability space.
+    N: (int)
+        Number of samples.
     s: (int)
         Thinning index
 
     """
-
-    p = q(r[0], r[1])  # Initial probability.
+    print("inits = ", r)
[…]

### 2017-02-12 chronometer/simple_gibbs.py
gibbs function

@@ -1,0 +1,173 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import emcee
+import corner
+
+
+def model(pars, fixed, x, fit):
+    """
+    The Model. A straight line plus a Gaussian.
+    params:
+    ------
+    pars: (tuple, list or string)
+        The parameters to vary.
+    fixed: (tuple, list or string)
+        The fixed parameters.
+    x: (array)
+        The x-array
+    fit: (int)
+        Either 0 or 1. If 0, the line parameters are read as pars and the
+        Gaussian parameters are read as fixed. The reverse is true if 1.
+    returns: (array)
+        The computed y-values of the model.
+    """
+    if fit == 0:
+        a, b = np.exp(pars)
+        mu, sig = np.exp(fixed)
+    elif fit == 1:
+        mu, sig = np.exp(pars)
+        a, b = np.exp(fixed)
+    return a + b*x + np.exp(-(x - mu)**2/(2.*sig**2))
+
+
+def line_lnprob(pars0, pars1, x, y, yerr):
+    """
+    The log-prob of the line model.
+    pars0: (tuple, list or string)
+        The line parameters, pars0 = [a, b]
+    pars1: (tuple, list or string)
+        The Gaussian parameters, pars1 = [mu, sigma]
+    x: (array)
+        The x-array
+    y: (array)
+        The observed y-values
+    yerr: (array)
+        The observed y-uncertainties.
+    returns: (float)
+        The log-probability.
+    """
+    mod = model(pars0, pars1, x, 0)
+    return sum(-((y - mod)**2/(2.*yerr**2))) + line_lnprior(pars0)
+
+
+def line_lnprior(pars0):
+    """
+    The log-prior for the parameters of the line model.
+    """
+    a, b = pars0
+    if -10 < a < 10 and -10 < b < 10:
+        return 0.
+    else:
+        return -np.inf
+
+
+def Gauss_lnprob(pars1, pars0, x, y, yerr):
+    """
+    The log-prob of the Gaussian model.
+    """
+    mod = model(pars1, pars0, x, 1)
+    return sum(-((y - mod)**2/(2.*yerr**2))) + Gauss_lnprior(pars1)
+
+
+def Gauss_lnprior(pars1):
+    """
+    The log-prior for the parameters of the Gaussian model.
+    """
+    mu, sig = pars1
+    if -10 < mu < 10 and -10 < sig < 10:
+        return 0.
+    else:
[…]

### 2017-02-13 chronometer/prototype.py
querying lnlike

@@ -1,0 +1,85 @@
+
+import os
+
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+import scipy.misc as spm
+from isochrones import StarModel
+from isochrones.mist import MIST_Isochrone
+
+
+def gc_model(data, params):
+    """
+    Given an age and a B-V colour, predict a rotation period.
+    Returns log(age) in Myr.
+    """
+    a, b, c, n = np.exp(params)
+    p, bv = data
+    return (np.log(p) - np.log(a) - b*np.log(bv - c))/n
+    # return (p/((bv - c)**b))**(1./n) * 1e-3
+
+
+def gc_lnlike(data, age, params, age_err):
+    """
+    Probability of rotation period and colour given age
+    """
+    model = gc_model(data, params)
+    return sum(-.5*((age - model)/age_err)**2)
+
+
+def gc_lnprior(params):
+    """
+    Some simple, uninformative prior over the parameters.
+    """
+    return 0.
+
+
+def iso_lnlike(data, age, age_err):
+    """
+    Some isochronal likelihood function.
+    """
+    logg, teff, feh = data
+    return 0.
+
+
+def iso_lnprior(params):
+    return 0.
+
+
+def lnprob(data, params, age, age_err):
+    logg, teff, feh, t, bv = data
+    return iso_lnlike([logg, teff, feh], age, age_err) + \
+        gc_lnlike([t, bv], params, age, age_err) + gc_lnprior(params) + \
+        iso_lnprior(params)
+
+
+if __name__ == "__main__":
+    mist = MIST_Isochrone()
+    mod = StarModel(mist, Teff=(5700, 100), logg=(4.5, 0.1), feh=(0.0, 0.1))
+    # pardict = {"Teff": 5700, "Teff_err": 100, "logg": 4.5, "logg_err": .1,
+    #            "feh": 0., "feh_err": .1}
+    p0 = [0.8, 9.5, 0.0, 200, 0.2]
+    print(mod.lnlike(p0))
+
+    assert 0
+    DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
+
+    periods = np.linspace(0, 100, 100)
+    bvs = np.ones(100) * .65
+    params = np.log([.7725, .601, .4, .5189])
+    loga = gc_model([periods, bvs], params)
+
+    # load data.
[…]

### 2017-02-13 chronometer/prototype.py
removing the c parameter

@@ -8,23 +8,25 @@ import scipy.misc as spm
 from isochrones import StarModel
 from isochrones.mist import MIST_Isochrone
 
+from simple_gibbs import gibbs
+import emcee
 
-def gc_model(data, params):
+
+def gc_model(params, data):
     """
     Given an age and a B-V colour, predict a rotation period.
     Returns log(age) in Myr.
     """
-    a, b, c, n = np.exp(params)
+    a, b, n = np.exp(params)
     p, bv = data
-    return (np.log(p) - np.log(a) - b*np.log(bv - c))/n
-    # return (p/((bv - c)**b))**(1./n) * 1e-3
+    return (np.log(p) - np.log(a) - b*np.log(bv - .4))/n
 
 
-def gc_lnlike(data, age, params, age_err):
+def gc_lnlike(params, data, age, age_err):
     """
     Probability of rotation period and colour given age
     """
-    model = gc_model(data, params)
+    model = gc_model(params, data)
     return sum(-.5*((age - model)/age_err)**2)
 
 
@@ -32,7 +34,15 @@ def gc_lnprior(params):
     """
     Some simple, uninformative prior over the parameters.
     """
-    return 0.
+    if (-10 < params[0] < 10) and (-10 < params[1] < 10) and \
+            (-10 < params[2] < 10):
+        return 0.
+    else:
+        return -np.inf
+
+
+def gc_lnprob(params, data, age, age_err):
+    return gc_lnlike(params, data, age, age_err) + gc_lnprior(params)
 
 
 def iso_lnlike(data, age, age_err):
@@ -55,31 +65,53 @@ def lnprob(data, params, age, age_err):
 
 
 if __name__ == "__main__":
-    mist = MIST_Isochrone()
-    mod = StarModel(mist, Teff=(5700, 100), logg=(4.5, 0.1), feh=(0.0, 0.1))
-    # pardict = {"Teff": 5700, "Teff_err": 100, "logg": 4.5, "logg_err": .1,
-    #            "feh": 0., "feh_err": .1}
-    p0 = [0.8, 9.5, 0.0, 200, 0.2]
-    print(mod.lnlike(p0))
 
-    assert 0
     DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
 
-    periods = np.linspace(0, 100, 100)
-    bvs = np.ones(100) * .65
[…]

### 2017-03-17 chronometer/chronometer.py
making more sense

@@ -1,3 +1,5 @@
+# Assembling likelihood functions for Gyrochronology, isochrones, dynamics
+# and asteroseismology.
 
 import os
 
@@ -10,61 +12,141 @@ from isochrones.mist import MIST_Isochrone
 
 from simple_gibbs import gibbs
 import emcee
+import corner
 
 
-def gc_model(params, data):
+def gc_model(params, bv):
     """
-    Given an age and a B-V colour, predict a rotation period.
+    Given a B-V colour and an age predict a rotation period.
     Returns log(age) in Myr.
+    parameters:
+    ----------
+    params: (array)
+        The array of age and (log) gyro parameters, a, b and n.
+    data: (array)
+        A an array containing colour.
     """
-    a, b, n = np.exp(params)
-    p, bv = data
-    return (np.log(p) - np.log(a) - b*np.log(bv - .4))/n
+    a, b, n, age = np.exp(params)
+    # return (np.log(p) - np.log(a) - b*np.log(bv - .4))/n
+    return n*np.log(age) + np.log(a) + b*np.log(bv - .4)
 
 
-def gc_lnlike(params, data, age, age_err):
+def gc_lnlike(params, data, data_err):
     """
-    Probability of rotation period and colour given age
+    Probability of age and model parameters given rotation period and colour.
+    parameters:
+    ----------
+    params: (array)
+        The array of log parameters: a, b, n, age.
+    data: (array)
+        an array containing rotation periods and B-V colour.
+    data_err: (array)
+        an array containing rotation period and B-V uncertainties.
     """
-    model = gc_model(params, data)
-    return sum(-.5*((age - model)/age_err)**2)
+    periods, bv = data
+    p_err, bv_err = data
+    model_periods = gc_model(params, bv)
+    return sum(-.5*((periods - model_periods)/p_err)**2)
 
 
-def gc_lnprior(params):
+def lnprior(params):
     """
     Some simple, uninformative prior over the parameters.
+    a, b, n, age, mass, feh, distance, Av = params
     """
-    if (-10 < params[0] < 10) and (-10 < params[1] < 10) and \
-            (-10 < params[2] < 10):
+    if -10 < params[0] < 10 and -10 < params[1] < 10 and \
[…]

### 2017-03-17 documents/ms.tex
words on the subway
testing the gyro model

@@
     Time variability & Time variability & ? & ? & ? \\
+    \hline
     Isochrones & $T_{\mathrm{eff}}$, $M_v$ & Core/shell fusion &
     Subgiants/giants & Physical (model dependent) \\
+    Spectroscopic twins & Spectra & ? & ? & ? \\
+    \hline
     Asteroseismology & Frequencies & Core/shell fusion & Old, low
     mass, Giants & Physical (model dependent) \\
-    White dwarf cooling & Cooling & $M_v$ & WDs (\& companions) &
+    \hline
+    White dwarf cooling & $M_v$ & Cooling & WDs (\& companions) &
     Physical (1 free parameter) \\
+    \hline
     $J_z$, $J_{\phi}$, $J_r$ & Position \& velocity & Dynamic heating & Disk
@@
     Myrs & 0 free parameters \\
-    Universe age & All of the above & Finite age of the Universe &
-    All stars & 0 free parameters \\
+    \hline
     C/N ratio & C/N ratio & Convective processes & Red giants, A \& F &
@@
     Myrs? & Effective \\
+    $[Y/Mg]$ & $[Y/Mg]$ & Enrichment & ? & Effective \\
+    \hline
     Nuclear isotopes & ? & ? & ? & ? \\
-    Unusual HRD objects & $T_{\mathrm{eff}}$, $M_v$ & Mergers, ? &
-    Special cases & ? \\
-    Twins & ? & ? & ? & ? \\
+    Unusual HRD objects & $T_{\mathrm{eff}}$, $M_v$ & Mergers, ? & Special
+    cases & ? \\
+    Universe age & All of the above & Finite age of the Universe &
+    All stars & 0 free parameters \\
 \end{tabular}
@@
 \racomment{What happens when you can't measure a rotation period?}
-\racomment{What about the difference in physics that you are probing?}
+\racomment{What about the difference in physics that you are investigating?}
 

### 2017-03-17 chronometer/action_age_evolution.py

@@ -9,6 +9,8 @@ from galpy.actionAngle import actionAngleStaeckel, actionAngleAdiabatic
 from galpy.df import quasiisothermaldf
 aA = actionAngleStaeckel(pot=MWPotential2014, delta=0.45, c=True)
 
+from mcquillan_granola import calc_bin_dispersion
+
 
 def calc_dispersion(pars, time):
     """
@@ -38,18 +40,30 @@ if __name__ == "__main__":
     sr = 34.
     sz = 25.1
 
-    zpars = [sz0, t1, beta, R0, Rc, hsz]
-    rpars = [sr0, t1, beta, R0, Rc, hsr]
+    zpars = [sz0, t1, beta, hsz]
+    rpars = [sr0, t1, beta, hsr]
 
     sigma_z = calc_dispersion(zpars, time)
     sigma_r = calc_dispersion(rpars, time)
     print(sigma_z)
 
+    # load data
+    DATA_DIR = "/Users/ruthangus/granola/granola/data"
+    d = pd.read_csv("ages_and_actions.csv")
+    m = (d.age.values > 0) * (d.age.values < 14)
+    df = d.iloc[m]
+
+    ages, dispersions, Ns = calc_bin_dispersion(df.age.values, df.Jz.values,
+                                                8)
+    d_err = dispersions / (2 * Ns - 2)**.5
+
     plt.clf()
     plt.plot(np.log(time), sigma_z)
+    plt.plot(ages, dispersions)
     plt.xlabel("log(Time)")
-    plt.ylabel("Sigma_z")
+    plt.ylabel("dispersion")
     plt.savefig("age_dispersion")
+
     assert 0
 
 

### 2017-03-17 chronometer/chronometer.py

@@ -27,8 +27,7 @@ def gc_model(params, bv):
         A an array containing colour.
     """
     a, b, n, age = np.exp(params)
-    # return (np.log(p) - np.log(a) - b*np.log(bv - .4))/n
-    return n*np.log(age) + np.log(a) + b*np.log(bv - .4)
+    return a*(age*1e3)**n * (bv[0] - .4)**b
 
 
 def gc_lnlike(params, data, data_err):
@@ -112,46 +111,37 @@ def lnprob(params, *args, fit_parallax=True, gyro=True,
         gc_lnprior(gyro_params) + iso_lnprior(iso_params)
 
 
-def fit_star():
-    bands = dict(
-            J=(star.jmag, star.jmag_err),
-            H=(star.hmag, star.hmag_err),
-            K=(star.kmag, star.kmag_err),
-        )
-    args = (bands, parallax, period, bv)
+def distance_modulus(M, D):
+    return 5*np.log10(D) - 5 + M
 
 
 if __name__ == "__main__":
 
-    # Testing the isochrones likelihood function.
-    mist = MIST_Isochrone()
-    mod = StarModel(mist, Teff=(5700, 100), logg=(4.5, 0.1), feh=(0.0, 0.1))
+    a, b, n = .7725, .601, .5189
+    age = 4.56
+    mass, feh, d, Av = 1., 0., 10., 0.
+    params = np.log(np.array([a, b, n, age, mass, feh, d, Av]))
 
-    import time
-    start = time.time()
-    # mass, age, metalicity, distance, extinction
-    p0 = [0.8, 9.5, 0.0, 200, 0.2]
-    print(mod.lnlike(p0))
-    end = time.time()
-    print(end - start, "seconds")
+    # test on the Sun first.
+    MJ = 3.711  # absolute magnitudes
+    MH = 3.453
+    MK = 3.357
+    J = distance_modulus(MJ, 10)
+    H = distance_modulus(MH, 10)
+    K = distance_modulus(MK, 10)
 
-    start = time.time()
-    p0 = [0.7, 9.5, 0.0, 200, 0.2]
-    print(mod.lnlike(p0))
-    end = time.time()
-    print(end - start, "seconds")
+    bands = dict(J=(J, .01), H=(H, .01), K=(K, .01),)
+    parallax = (.1, .001)
+    period = (26., 1.)
+    bv = (.65, .01)
+    args = (bands, parallax, period, bv)
 
-    start = time.time()
[…]

### 2017-03-17 chronometer/simple_gibbs.py

@@ -145,7 +145,6 @@ def gibbs(par_init0, par_init1, x, y, yerr, niter):
         print("Model {0}, Iteration {1}".format(1, i))
         p1, flat1 = gibbs_internal(p0, p1, x, y, yerr, 1)
 
-    print(np.shape(flat0), np.shape(flat1))
     flat = np.concatenate((flat0, flat1), axis=1)
     fig = corner.corner(flat)
     fig.savefig("corner")

### 2017-03-17 chronometer/staying_alive.py

@@ -1,146 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.mlab as mlab
-import scipy.misc as spm
-import corner
-
-
-#!/usr/bin/env python
-# Author:      Flávio Codeço Coelho
-# License:     GPL
-
-from math import *
-import numpy as np
-from matplotlib.pylab import *
-
-# def sdnorm(z):
-#     """
-#     Standard normal pdf (Probability Density Function)
-#     """
-#     return exp(-z*z/2.)/sqrt(2*pi)
-
-# n = 10000
-# alpha = 1
-# x = 0.
-# vec = []
-# vec.append(x)
-# innov = np.random.uniform(-alpha,alpha,n) #random inovation, uniform proposal distribution
-# for i in range(n):
-#     can = x + innov[i] #candidate
-#     aprob = min([1.,sdnorm(can)/sdnorm(x)]) #acceptance probability
-#     u = uniform(0,1)
-#     if u < aprob:
-#         x = can
-#         vec.append(x)
-
-# #plotting the results:
-# #theoretical curve
-# x = arange(-3,3,.1)
-# y = sdnorm(x)
-# subplot(211)
-# title('Metropolis-Hastings')
-# plot(vec)
-# subplot(212)
-
-# hist(vec, bins=30,normed=1)
-# plot(x,y,'ro')
-# ylabel('Frequency')
-# xlabel('x')
-# legend(('PDF','Samples'))
-# show()
-
-def q(x, y):
-    g1 = mlab.bivariate_normal(x, y, 1.0, 1.0, -1, -1, -0.8)
-    g2 = mlab.bivariate_normal(x, y, 1.5, 0.8, 1, 2, 0.6)
-    return 0.6*g1+28.4*g2/(0.6+28.4)
-
-
-def model(par, x):
-    return par[0] + par[1]*x
-
-
-def lnlike(par, x, y_obs, y_err):
-    y_mod = model(par, x)
-    return sum(-.5*((y_mod - y_obs)/y_err)**2)
-
-
-def metropolis_hastings(x, y, yerr, r, N, s, t):
-    """
-    params:
-    -------
-    r: (tuple or list or array)
-        The parameters/position in probability space.
-    N: (int)
-        Number of samples.
-    s: (int)
-        Thinning index
-
-    """
-    print("inits = ", r)
-    p = np.exp(lnlike(r, x, y, yerr))  # Initial probability.
-    samples = []
-    for i in np.arange(N):
[…]

### 2017-03-17 chronometer/chronometer.py
non-zero iso_lnlike

@@ -3,6 +3,7 @@
 
 import os
 
+import time
 import numpy as np
 import matplotlib.pyplot as plt
 import pandas as pd
@@ -27,25 +28,40 @@ def gc_model(params, bv):
         A an array containing colour.
     """
     a, b, n, age = np.exp(params)
-    return a*(age*1e3)**n * (bv[0] - .4)**b
+    return a*(age*1e3)**n * (bv - .4)**b
 
 
-def gc_lnlike(params, data, data_err):
+def gc_lnlike(params, period, bv):
     """
     Probability of age and model parameters given rotation period and colour.
     parameters:
     ----------
     params: (array)
         The array of log parameters: a, b, n, age.
-    data: (array)
-        an array containing rotation periods and B-V colour.
-    data_err: (array)
-        an array containing rotation period and B-V uncertainties.
+    period: (tuple)
+        The rotation period and period uncertainty in days
+    bv: (tuple)
+        The B-V colour and colour uncertainty.
     """
-    periods, bv = data
-    p_err, bv_err = data
-    model_periods = gc_model(params, bv)
-    return sum(-.5*((periods - model_periods)/p_err)**2)
+    model_periods = gc_model(params, bv[0])
+    return sum(-.5*((period[0] - model_periods)/period[1])**2)
+
+
+def iso_lnlike(lnparams):
+    """
+    Some isochronal likelihood function.
+    parameters:
+    ----------
+    params: (array)
+        The array of parameters: mass, age, metallicity, distance, extinction.
+    bands: (dict)
+        The dictionary of Kepler j, h, k with their uncertainties.
+    parallax: (tuple)
+        The parallax and its uncertainty.
+    """
+    params = np.array([np.exp(lnparams[0]), np.log10(1e9*np.exp(lnparams[1])),
+                      lnparams[2], np.exp(lnparams[3]), lnparams[4]])
+    return mod.lnlike(params)
 
 
 def lnprior(params):
@@ -60,29 +76,6 @@ def lnprior(params):
         return 0.
     else:
         return -np.inf
-
-
-def iso_lnlike(params, bands, parallax, fit_parallax=True):
-    """
-    Some isochronal likelihood function.
-    parameters:
-    ----------
-    params: (array)
[…]

### 2017-03-17 chronometer/chronometer.py
iso_lnlike speedup

@@ -47,21 +47,21 @@ def gc_lnlike(params, period, bv):
     return sum(-.5*((period[0] - model_periods)/period[1])**2)
 
 
-def iso_lnlike(lnparams):
+def iso_lnlike(lnparams, mods):
     """
     Some isochronal likelihood function.
     parameters:
     ----------
     params: (array)
-        The array of parameters: mass, age, metallicity, distance, extinction.
-    bands: (dict)
-        The dictionary of Kepler j, h, k with their uncertainties.
-    parallax: (tuple)
-        The parallax and its uncertainty.
+        The array of parameters: log(mass), log(age in Gyr), metallicity,
+        log(distance), extinction.
+    mods: (list)
+        A list of isochrones.py starmodel objects.
     """
     params = np.array([np.exp(lnparams[0]), np.log10(1e9*np.exp(lnparams[1])),
                       lnparams[2], np.exp(lnparams[3]), lnparams[4]])
-    return mod.lnlike(params)
+    ll = [mod.lnlike(params) for mod in mods]
+    return sum(ll)
 
 
 def lnprior(params):
@@ -78,30 +78,22 @@ def lnprior(params):
         return -np.inf
 
 
-def lnprob(params, *args, fit_parallax=True, gyro=True,
-           iso=True):
+def lnprob(params, mods, *args, gyro=True, iso=True):
     """
     The joint log-probability of age given gyro and iso parameters.
-    bands: (dict)
-        dictionary of Kepler bands and their uncertainties.
-    parallax: (tuple)
-        parallax and its uncertainty, etc.
+    mods: (list)
+        list of pre-computed star model objects.
+    gyro: (bool)
+        If True, the gyro likelihood will be used.
+    iso: (bool)
+        If True, the iso likelihood will be used.
     """
-    if fit_parallax:
-        bands, parallax, period, bv = args
-        isoargs = (bands, parallax)
-    else:
-        bands, period, bv = args
-        isoargs = bands
-    gyroargs = (period, bv)
-
     a, b, n, age, mass, feh, d, Av = params
     iso_params = np.array([mass, age, feh, d, Av])
     gyro_params = np.array([a, b, n, age])
 
-    return iso_lnlike(iso_params, *isoargs) + \
-        gc_lnlike(gyro_params, gyro_args) + \
[…]

### 2017-03-17 chronometer/chronometer.py
testing lnprob

@@ -92,8 +92,8 @@ def lnprob(params, mods, *args, gyro=True, iso=True):
     iso_params = np.array([mass, age, feh, d, Av])
     gyro_params = np.array([a, b, n, age])
 
-    return iso_lnlike(iso_params, mods) + gc_lnlike(gyro_params, args) + \
-        lnprior(params)
+    return iso_lnlike(iso_params, mods) + gc_lnlike(gyro_params, period, bv) \
+        + lnprior(params)
 
 
 def distance_modulus(M, D):
@@ -105,7 +105,8 @@ if __name__ == "__main__":
     a, b, n = .7725, .601, .5189
     age = 4.56
     mass, feh, d, Av = 1., 0., 10., 0.
-    params = np.log(np.array([a, b, n, age, mass, feh, d, Av]))
+    params = np.array([a, b, n, np.log(age), np.log(mass), feh, np.log(d),
+                       Av])
 
     N = 10
     _1s = np.ones(N)  # make fake data arrays, length 10.
@@ -126,7 +127,7 @@ if __name__ == "__main__":
     # print(gc_model(gyro_params, bv[0]))
 
     # test the gyro lhf
-    # print(gc_lnlike(gyro_params, period, bv))
+    print("gyro_lnlike = ", gc_lnlike(gyro_params, period, bv))
 
     # test the iso_lnlike
     mist = MIST_Isochrone()
@@ -147,11 +148,12 @@ if __name__ == "__main__":
     print("preamble time = ", end - start)
 
     start = time.time()
-    print(iso_lnlike(iso_params, mods))
+    print("iso_lnlike = ", iso_lnlike(iso_params, mods))
     end = time.time()
     print("lhf time = ", end - start)
 
     # test the lnprob.
+    print("lnprob = ", lnprob(params, mods, period, bv, gyro=True, iso=True))
 
 
     DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"

### 2017-03-18 chronometer/chronometer.py
getting started with one star

@@ -27,8 +27,8 @@ def gc_model(params, bv):
     data: (array)
         A an array containing colour.
     """
-    a, b, n, age = np.exp(params)
-    return a*(age*1e3)**n * (bv - .4)**b
+    a, b, n, ln_age = params
+    return a*(np.exp(ln_age)*1e3)**n * (bv - .4)**b
 
 
 def gc_lnlike(params, period, bv):
@@ -44,10 +44,10 @@ def gc_lnlike(params, period, bv):
         The B-V colour and colour uncertainty.
     """
     model_periods = gc_model(params, bv[0])
-    return sum(-.5*((period[0] - model_periods)/period[1])**2)
+    return -.5*((period[0] - model_periods)/period[1])**2
 
 
-def iso_lnlike(lnparams, mods):
+def iso_lnlike(lnparams, mod):
     """
     Some isochronal likelihood function.
     parameters:
@@ -55,13 +55,12 @@ def iso_lnlike(lnparams, mods):
     params: (array)
         The array of parameters: log(mass), log(age in Gyr), metallicity,
         log(distance), extinction.
-    mods: (list)
-        A list of isochrones.py starmodel objects.
+    mod: (object)
+        An isochrones.py starmodel object.
     """
     params = np.array([np.exp(lnparams[0]), np.log10(1e9*np.exp(lnparams[1])),
                       lnparams[2], np.exp(lnparams[3]), lnparams[4]])
-    ll = [mod.lnlike(params) for mod in mods]
-    return sum(ll)
+    return mod.lnlike(params)
 
 
 def lnprior(params):
@@ -78,10 +77,10 @@ def lnprior(params):
         return -np.inf
 
 
-def lnprob(params, mods, *args, gyro=True, iso=True):
+def lnprob(params, mod, period, bv, gyro=True, iso=False):
     """
     The joint log-probability of age given gyro and iso parameters.
-    mods: (list)
+    mod: (list)
         list of pre-computed star model objects.
     gyro: (bool)
         If True, the gyro likelihood will be used.
@@ -92,8 +91,13 @@ def lnprob(params, mods, *args, gyro=True, iso=True):
     iso_params = np.array([mass, age, feh, d, Av])
     gyro_params = np.array([a, b, n, age])
 
-    return iso_lnlike(iso_params, mods) + gc_lnlike(gyro_params, period, bv) \
[…]

### 2017-03-19 chronometer/iso_like_test.py
adding iso_prob_test

@@ -1,0 +1,131 @@
+# Assembling likelihood functions for Gyrochronology, isochrones, dynamics
+# and asteroseismology.
+
+import matplotlib.pyplot as plt
+import time
+import numpy as np
+from isochrones import StarModel
+from isochrones.mist import MIST_Isochrone
+
+import emcee
+import corner
+
+
+def iso_lnlike(lnparams, mod):
+    """
+    Some isochronal likelihood function.
+    parameters:
+    ----------
+    params: (array)
+        The array of parameters: log(mass), log(age in Gyr), metallicity,
+        log(distance), extinction.
+    mod: (object)
+        An isochrones.py starmodel object.
+    """
+    lnmass, lnage, feh, lnd, Av = lnparams
+    params = np.array([np.exp(lnmass), np.log10(1e9*np.exp(lnage)), feh,
+                       np.exp(lnd), Av])
+    return mod.lnlike(params)
+
+
+def lnprior(params):
+    """
+    Some simple, uninformative prior over the parameters.
+    age, mass, feh, distance, Av = params
+    """
+    if -10 < params[0] < 11 and -10 < params[1] < 10 and \
+            -10 < params[2] < 10 and -10 < params[3] < 11 and \
+            -10 < params[4] < 10:
+        return 0.
+    else:
+        return -np.inf
+
+
+def lnprob(params, mod):
+    """
+    The joint log-probability of age given gyro and iso parameters.
+    mod: (list)
+        list of pre-computed star model objects.
+    """
+
+    return iso_lnlike(params, mod) + lnprior(params)
+
+
+def probtest(xs, i):
+    lps = []
+    for x in xs:
+        params[i] = x
+        lp = lnprob(params, mod)
+        lps.append(lp)
+    plt.clf()
+    plt.plot(xs, lps)
+    plt.xlabel("X")
+    plt.ylabel("lnprob")
+    plt.savefig("probs")
+
+
+if __name__ == "__main__":
+
+    age = 4.56
+    mass, feh, d, Av = 1., 0., 10., 0.
+    params = np.array([np.log(mass), np.log(age), feh, np.log(d), Av])
+
+    # test on the Sun at 10 pc first.
+    J, J_err = 3.711, .01  # absolute magnitudes/apparent at D = 10pc
[…]

### 2017-03-19 chronometer/chronometer.py
parameterisation bug

@@ -43,7 +43,7 @@ def gc_lnlike(params, period, bv):
     bv: (tuple)
         The B-V colour and colour uncertainty.
     """
-    model_periods = gc_model(params, bv[0])
+    model_periods = gc_model(params[:4], bv[0])
     return -.5*((period[0] - model_periods)/period[1])**2
 
 
@@ -58,8 +58,9 @@ def iso_lnlike(lnparams, mod):
     mod: (object)
         An isochrones.py starmodel object.
     """
-    params = np.array([np.exp(lnparams[0]), np.log10(1e9*np.exp(lnparams[1])),
-                      lnparams[2], np.exp(lnparams[3]), lnparams[4]])
+    a, b, n, lnage, lnmass, feh, lnd, Av = lnparams
+    params = np.array([np.exp(lnmass), np.log10(1e9*np.exp(lnage)), feh,
+                       np.exp(lnd), Av])
     return mod.lnlike(params)
 
 
@@ -69,7 +70,7 @@ def lnprior(params):
     a, b, n, age, mass, feh, distance, Av = params
     """
     if -10 < params[0] < 10 and -10 < params[1] < 10 and \
-            -10 < params[2] < 10 and -10 < params[3] < 10 and \
+            -10 < params[2] < 10 and -10 < params[3] < 11 and \
             -10 < params[4] < 10 and -10 < params[5] < 10 and \
             -10 < params[6] < 10 and -10 < params[7] < 10:
         return 0.
@@ -77,7 +78,7 @@ def lnprior(params):
         return -np.inf
 
 
-def lnprob(params, mod, period, bv, gyro=True, iso=False):
+def lnprob(params, mod, period, bv, gyro=False, iso=True):
     """
     The joint log-probability of age given gyro and iso parameters.
     mod: (list)
@@ -87,21 +88,28 @@ def lnprob(params, mod, period, bv, gyro=True, iso=False):
     iso: (bool)
         If True, the iso likelihood will be used.
     """
-    a, b, n, age, mass, feh, d, Av = params
-    iso_params = np.array([mass, age, feh, d, Av])
-    gyro_params = np.array([a, b, n, age])
 
     if gyro and iso:
-        return iso_lnlike(iso_params, mod) + gc_lnlike(gyro_params, period,
-                                                       bv) + lnprior(params)
[…]

### 2017-03-19 chronometer/chronometer.py
removing cluster data

@@ -187,13 +187,3 @@ if __name__ == "__main__":
               "$\ln(D)$", "$A_v$"]
     fig = corner.corner(flat, labels=labels)
     fig.savefig("corner_test")
-
-    DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
-    # load data.
-    cs = pd.read_csv(os.path.join(DATA_DIR, "clusters.csv"))
-    m = (.6 < cs.bv.values) * (cs.bv.values < .7)
-    cs = cs.iloc[m]
-    periods = cs.period.values
-    bvs = cs.bv.values
-    ages = cs.age.values * 1e3
-    age_err = cs.age_err.values

### 2017-03-19 chronometer/chronometer.py
still looking for the bug...

@@ -145,7 +145,7 @@ if __name__ == "__main__":
     # iso_lnlike preamble.
     start = time.time()
     mod = StarModel(mist, J=(J, J_err), H=(H, H_err), K=(K, K_err),
-                    parallax=(.1, .001))
+                    parallax=(.1, .001), Teff=(5700, 50), logg=(4.44, .1))
     p0 = np.array([params[0], params[1], params[2], params[3], params[4]])
 
     start = time.time()
@@ -178,7 +178,7 @@ if __name__ == "__main__":
     sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob,
                                     args=[mod, period, bv])
     print("burning in...")
-    pos, _, _ = sampler.run_mcmc(p0, 1000)
+    pos, _, _ = sampler.run_mcmc(p0, 10)
     sampler.reset()
     print("production run...")
     sampler.run_mcmc(pos, nsteps)
@@ -187,3 +187,9 @@ if __name__ == "__main__":
               "$\ln(D)$", "$A_v$"]
     fig = corner.corner(flat, labels=labels)
     fig.savefig("corner_test")
+
+    for i in range(ndim):
+        plt.clf()
+        plt.plot(flat[:, i], alpha=.5)
+        plt.ylabel(labels[i])
+        plt.savefig("{}_trace".format(i))

### 2017-03-19 chronometer/iso_like_test.py

@@ -53,9 +53,10 @@ def lnprob(params, mod):
 
 def probtest(xs, i):
     lps = []
+    p = params + 0
     for x in xs:
-        params[i] = x
-        lp = lnprob(params, mod)
+        p[i] = x
+        lp = lnprob(p, mod)
         lps.append(lp)
     plt.clf()
     plt.plot(xs, lps)
@@ -115,12 +116,12 @@ if __name__ == "__main__":
     # assert 0
 
     # Run MCMC
-    nwalkers, nsteps, ndim = 64, 10000, len(params)
+    nwalkers, nsteps, ndim = 64, 1000, len(params)
     p0 = [1e-4*np.random.rand(ndim) + params for i in range(nwalkers)]
     sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob, args=[mod])
 
     print("burning in...")
-    pos, _, _ = sampler.run_mcmc(p0, 1000)
+    pos, _, _ = sampler.run_mcmc(p0, 10)
     sampler.reset()
 
     print("production run...")
@@ -129,3 +130,10 @@ if __name__ == "__main__":
     labels = ["$Age$", "$Mass$", "$[Fe/H]$", "$D$", "$A_v$"]
     fig = corner.corner(flat)
     fig.savefig("corner_iso")
+
+    print(np.shape(flat))
+    for i in range(ndim):
+        plt.clf()
+        plt.plot(flat[:, i], "k", alpha=.5)
+        plt.ylabel(labels[i])
+        plt.savefig("{}_trace".format(i))

### 2017-03-19 chronometer/priors.py
adding priors from tdm isochrones.py

@@ -1,0 +1,69 @@
+import numpy as np
+
+def age_prior(age, bounds=(9,10.15)):
+    """
+    Uniform true age prior; where 'age' is actually log(age)
+    """
+    minage, maxage = bounds
+    if age < minage or age > maxage:
+        return 0
+    return np.log(10) * 10**age / (10**maxage - 10**minage)
+
+def distance_prior(distance, bounds=(0,3000)):
+    """
+    Distance prior ~ d^2
+    """
+    min_distance, max_distance = bounds
+    if distance <= min_distance or distance > max_distance:
+        return 0
+    return 3/max_distance**3 * distance**2
+
+def AV_prior(AV, bounds=(0,1.)):
+    if AV < bounds[0] or AV > bounds[1]:
+        return 0
+    return 1./bounds[1]
+
+def q_prior(q, m=1, gamma=0.3, bounds=(0.1,1)):
+    """Default prior on mass ratio q ~ q^gamma
+    """
+    qmin, qmax = bounds
+    try:
+        if q < qmin or q > qmax:
+            return 0
+    except ValueError:
+        pass
+
+    C = 1/(1/(gamma+1)*(1 - qmin**(gamma+1)))
+    result = C*q**gamma
+    if np.size(result) > 1:
+        result[(q < qmin) | (q > qmax)] = 0
+    return result
+
+def salpeter_prior(m,alpha=-2.35, bounds=(0.1,10)):
+    minmass, maxmass = bounds
+    C = (1+alpha)/(maxmass**(1+alpha)-minmass**(1+alpha))
+    try:
+        if m < minmass or m > maxmass:
+            return 0
+        else:
+            return C*m**(alpha)
+    except ValueError:
+        result = C*m**(alpha)
+        result[(m < minmass) | (m > maxmass)] = 0
+        return result
+
+def feh_prior(feh, bounds=None, halo_fraction=0.001):
+    """feh PDF based on local SDSS distribution
+
+    From Jo Bovy:
+    https://github.com/jobovy/apogee/blob/master/apogee/util/__init__.py#L3
+    2D gaussian fit based on Casagrande (2011)
+    """
+    disk_norm = 2.5066282746310007 # integral of the below from -np.inf to np.inf
[…]

### 2017-03-19 chronometer/chronometer.py
adding priors

@@ -14,6 +14,7 @@ from isochrones.mist import MIST_Isochrone
 from simple_gibbs import gibbs
 import emcee
 import corner
+import priors
 
 
 def gc_model(params, bv):
@@ -65,15 +66,13 @@ def iso_lnlike(lnparams, mod):
 
 
 def lnprior(params):
-    """
-    Some simple, uninformative prior over the parameters.
-    a, b, n, age, mass, feh, distance, Av = params
-    """
+    age_prior = np.log(priors.age_prior(np.log10(1e9*np.exp(params[3]))))
+    feh_prior = np.log(priors.feh_prior(params[5]))
+    distance_prior = np.log(priors.distance_prior(np.exp(params[6])))
+    Av_prior = np.log(priors.AV_prior(params[7]))
     if -10 < params[0] < 10 and -10 < params[1] < 10 and \
-            -10 < params[2] < 10 and -10 < params[3] < 11 and \
-            -10 < params[4] < 10 and -10 < params[5] < 10 and \
-            -10 < params[6] < 10 and -10 < params[7] < 10:
-        return 0.
+            -10 < params[2] < 10:
+        return age_prior + feh_prior + distance_prior + Av_prior
     else:
         return -np.inf
 
@@ -178,7 +177,7 @@ if __name__ == "__main__":
     sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob,
                                     args=[mod, period, bv])
     print("burning in...")
-    pos, _, _ = sampler.run_mcmc(p0, 10)
+    pos, _, _ = sampler.run_mcmc(p0, 1000)
     sampler.reset()
     print("production run...")
     sampler.run_mcmc(pos, nsteps)
@@ -188,8 +187,13 @@ if __name__ == "__main__":
     fig = corner.corner(flat, labels=labels)
     fig.savefig("corner_test")
 
+    plt.clf()
+    print(np.shape(sampler.lnprobability.T))
+    plt.plot(sampler.lnprobability.T, "k")
+    plt.savefig("prob_trace")
+
     for i in range(ndim):
         plt.clf()
-        plt.plot(flat[:, i], alpha=.5)
+        plt.plot(sampler.chain[:, :,  i].T, alpha=.5)
         plt.ylabel(labels[i])
         plt.savefig("{}_trace".format(i))

### 2017-03-20 chronometer/chronometer.py
I think it's working. Now adding multiple stars.

@@ -77,7 +77,7 @@ def lnprior(params):
         return -np.inf
 
 
-def lnprob(params, mod, period, bv, gyro=False, iso=True):
+def lnprob(params, mod, period, bv, gyro=True, iso=True):
     """
     The joint log-probability of age given gyro and iso parameters.
     mod: (list)
@@ -144,7 +144,7 @@ if __name__ == "__main__":
     # iso_lnlike preamble.
     start = time.time()
     mod = StarModel(mist, J=(J, J_err), H=(H, H_err), K=(K, K_err),
-                    parallax=(.1, .001), Teff=(5700, 50), logg=(4.44, .1))
+                    parallax=(.1, .001))
     p0 = np.array([params[0], params[1], params[2], params[3], params[4]])
 
     start = time.time()
@@ -158,7 +158,7 @@ if __name__ == "__main__":
     print("lhf time = ", end - start)
 
     # test the lnprob.
-    print("lnprob = ", lnprob(params, mod, period, bv, gyro=False, iso=True))
+    print("lnprob = ", lnprob(params, mod, period, bv, gyro=True, iso=True))
 
     ages = np.log(np.arange(1., 10., 1))
     masses = np.log(np.arange(.1, 2., .1))
@@ -182,9 +182,11 @@ if __name__ == "__main__":
     print("production run...")
     sampler.run_mcmc(pos, nsteps)
     flat = np.reshape(sampler.chain, (nwalkers*nsteps, ndim))
+    truths = [.7725, .601, .5189, np.log(4.56), np.log(1), 0., np.log(10),
+              0.]
     labels = ["$a$", "$b$", "$n$", "$\ln(Age)$", "$\ln(Mass)$", "$[Fe/H]$",
               "$\ln(D)$", "$A_v$"]
-    fig = corner.corner(flat, labels=labels)
+    fig = corner.corner(flat, labels=labels, truths=truths)
     fig.savefig("corner_test")
 
     plt.clf()

### 2017-03-20 chronometer/single_chronometer.py
adding more stars

@@ -119,6 +119,7 @@ if __name__ == "__main__":
     params = np.array([a, b, n, np.log(age), np.log(mass), feh, np.log(d),
                        Av])
 
+
     # test on the Sun at 10 pc first.
     J, J_err = 3.711, .01  # absolute magnitudes/apparent at D = 10pc
     H, H_err = 3.453, .01

### 2017-05-01 chronometer/chronometer.py
nsteps is now a list

@@ -51,7 +51,6 @@ def lnlike(params, *args):
     N = len(period)
     gyro_lnlike, iso_lnlike = 0, 0
     if par_inds[0] == 0 and par_inds[1] == 1 and par_inds[2] == 2: # if gyro.
-
         gyro_lnlike = sum(-.5*((period - gyro_model(params, bv))
                                 /period_errs)**2)
 
@@ -152,7 +151,8 @@ def MH(par, lnprob, nsteps, t, *args):
         accept += acc
         probs.append(new_prob)
         samples[i, :] = par[par_inds]
-    print("Acceptance fraction = ", accept/float(nsteps))
+    if nsteps > 0:
+        print("Acceptance fraction = ", accept/float(nsteps))
     return samples, par, probs
 
 
@@ -188,7 +188,7 @@ def MH_step(par, lnprob, t, *args, emc=False):
     return par, new_lnprob, accept
 
 
-def gibbs_control(par, lnprob, nsteps, niter, t, par_inds_list, args):
+def gibbs_control(par, lnprob, nsteps_list, niter, t, par_inds_list, args):
     """
     This function tells the metropolis hastings what parameters to sample in.
     params:
@@ -222,7 +222,7 @@ def gibbs_control(par, lnprob, nsteps, niter, t, par_inds_list, args):
     n_parameter_sets = len(par_inds_list)
 
     # Final sample array
-    all_samples = np.zeros((nsteps * niter, ndim + nstars))
+    all_samples = np.zeros((sum(nsteps_list) * niter, ndim + nstars))
 
     # Iterate over niter cycles of parameter sets.
     probs = []
@@ -230,6 +230,8 @@ def gibbs_control(par, lnprob, nsteps, niter, t, par_inds_list, args):
         print("Gibbs iteration ", i, "of ", niter)
         print(par)
         for k in range(len(par_inds_list)):  # loop over parameter sets.
+            nsteps = nsteps_list[i]
+            print(nsteps, "steps")
             args[-1] = par_inds_list[k]
             samples, par, pb = MH(par, lnprob, nsteps, t[k], *args)
             if par_inds_list[k][0] == 0:  # save age samples separately
@@ -282,7 +284,9 @@ if __name__ == "__main__":
 
[…]

### 2017-05-01 chronometer/chronometer.py
fixing bug

@@ -285,7 +285,7 @@ if __name__ == "__main__":
     start = time.time()  # timeit
 
     # Different nsteps for different parameters. gyro, star 1, 2, 3.
-    nsteps = [0, 100, 0, 0]
+    nsteps = [0, 10000, 0, 0]
     niter = 2
     N = len(mods)
 
@@ -348,9 +348,9 @@ if __name__ == "__main__":
     print("Plotting results and traces")
     plt.clf()
     if run_MH:
-        for j in range(nsteps*niter - 1):
-                x = np.arange(j*nsteps[i], (j+1)*nsteps[i])
-                plt.plot(x, lnprobs[j*nsteps[i]: (j+1)*nsteps[i]])
+        for j in range(nsteps[1]*niter - 1):
+                x = np.arange(j*nsteps[1], (j+1)*nsteps[1])
+                plt.plot(x, lnprobs[j*nsteps[1]: (j+1)*nsteps[1]])
     else:
         plt.plot(sampler.lnprobability.T)
     plt.xlabel("Time")

### 2017-05-01 chronometer/chronometer.py
wrong index bug fix

@@ -230,8 +230,8 @@ def gibbs_control(par, lnprob, nsteps_list, niter, t, par_inds_list, args):
         print("Gibbs iteration ", i, "of ", niter)
         print(par)
         for k in range(len(par_inds_list)):  # loop over parameter sets.
-            nsteps = nsteps_list[i]
-            print(nsteps, "steps")
+            nsteps = nsteps_list[k]
+            print(nsteps, "steps", "parameter_indices = ", par_inds_list[k])
             args[-1] = par_inds_list[k]
             samples, par, pb = MH(par, lnprob, nsteps, t[k], *args)
             if par_inds_list[k][0] == 0:  # save age samples separately
@@ -285,7 +285,7 @@ if __name__ == "__main__":
     start = time.time()  # timeit
 
     # Different nsteps for different parameters. gyro, star 1, 2, 3.
-    nsteps = [0, 10000, 0, 0]
+    nsteps = [0, 100000, 0, 0]
     niter = 2
     N = len(mods)
 
@@ -345,17 +345,17 @@ if __name__ == "__main__":
     end = time.time()
     print("Time taken = ", (end - start)/60, "minutes")
 
-    print("Plotting results and traces")
-    plt.clf()
-    if run_MH:
-        for j in range(nsteps[1]*niter - 1):
-                x = np.arange(j*nsteps[1], (j+1)*nsteps[1])
-                plt.plot(x, lnprobs[j*nsteps[1]: (j+1)*nsteps[1]])
-    else:
-        plt.plot(sampler.lnprobability.T)
-    plt.xlabel("Time")
-    plt.ylabel("ln (probability)")
-    plt.savefig(os.path.join(RESULTS_DIR, "prob_trace"))
+    # print("Plotting results and traces")
+    # plt.clf()
+    # if run_MH:
+    #     for j in range(nsteps[1]*niter - 1):
+    #             x = np.arange(j*nsteps[1], (j+1)*nsteps[1])
+    #             plt.plot(x, lnprobs[j*nsteps[1]: (j+1)*nsteps[1]])
+    # else:
+    #     plt.plot(sampler.lnprobability.T)
+    # plt.xlabel("Time")
+    # plt.ylabel("ln (probability)")
[…]

### 2017-05-01 chronometer/mini_steps_chronometer.py
changing proposal

@@ -220,6 +220,15 @@ def estimate_covariance():
     return np.cov(samples, rowvar=False)
 
 
+def estimate_covariance():
+    """
+    Return the covariance matrix of the emcee samples.
+    """
+    with h5py.File("emcee_posterior_samples.h5", "r") as f:
+        samples = f["samples"][...]
+    return np.cov(samples, rowvar=False)
+
+
 if __name__ == "__main__":
 
     # Use Metropolis hastings or emcee?
@@ -236,12 +245,13 @@ if __name__ == "__main__":
 
     start = time.time()  # timeit
 
-    nsteps, niter = 1000, 10
+    nsteps, niter = 1000, 5
 
     # Construct parameter indices for the different parameter sets.
     par_inds_list = np.arange(len(params))
 
-    t = np.ones(len(params)) * 1e-1
+    t = estimate_covariance()[0]
+    # t = np.ones(len(params)) * 1e-1
     # t = np.array([7e-2, 7e-2, 7e-2, 1e-5, 1e-5, 1e-5, 1e-2, 1e-2, 1e-2,
     #               1e-1, 1e-3, 1e-3, 1e-3, 1e-3, 1e-3, 1e-3, 1e-3, 1e-3])
 
@@ -277,7 +287,7 @@ if __name__ == "__main__":
                 "$A_{v2}$", "$A_{v3}$"]
     print(np.shape(flat), "Sample shape")
     fig = corner.corner(flat, truths=truths, labels=labels)
-    fig.savefig(os.path.join(RESULTS_DIR, "demo_corner_gibbs"))
+    fig.savefig(os.path.join(RESULTS_DIR, "mini_corner_gibbs"))
 
     print("Plotting chains...")
     ndim = len(params)

### 2017-05-01 chronometer/MH_single.py
testing against chronometer

@@ -113,7 +113,7 @@ def MH(par, lnprob, nsteps, t, mod, emc=False):
     return samples, par, probs
 
 
-def MH_step(par, lnprob, t, *args, emc=True):
+def MH_step(par, lnprob, t, *args, emc=False):
     if emc:
         nwalkers, ndim = 10, len(par)
         p0 = [par + np.random.multivariate_normal(np.zeros((len(par))), t)

### 2017-05-02 chronometer/MH_single.py
diagnosing bug

@@ -40,6 +40,7 @@ def lnlike(params, mod):
     par_inds: (array)
         The parameters to vary.
     """
+    print(params, mod)
     p = params*1
     p[0] = np.exp(p[0])
     p[1] = np.log10(1e9*np.exp(p[1]))
@@ -54,9 +55,10 @@ def lnprior(params):
     age_prior = np.log(priors.age_prior(np.log10(1e9*np.exp(params[1]))))
     feh_prior = np.log(priors.feh_prior(params[2]))
     distance_prior = np.log(priors.distance_prior(np.exp(params[3])))
+
     m = (-20 < params) * (params < 20)  # Broad bounds on all params.
-    Av = params[4]
-    mAv = (0 <= Av) * (Av < 1)
+    mAv = (0 <= params[4]) * (params[4] < 1)
+
     if sum(m) == len(m) and mAv:
         return age_prior + feh_prior + distance_prior
     else:
@@ -71,6 +73,8 @@ def lnprob(params, mod):
     mod: (list)
         list of pre-computed star model objects.
     """
+    print(lnlike(params, mod), lnprior(params))
+    input("e")
     return lnlike(params, mod) + lnprior(params)
 
 
@@ -150,7 +154,10 @@ if __name__ == "__main__":
     par_inds = np.array([3, 6, 9, 12, 15])
     mods = mods[0]
 
-    nsteps = 1000
+    lnprob(params, mods)
+    input('e')
+
+    nsteps = 10000
 
     with h5py.File("emcee_posterior_samples.h5", "r") as f:
         samples = f["samples"][...]

### 2017-05-02 chronometer/chronometer.py

@@ -84,23 +84,33 @@ def emcee_lnprob(params, *args):
     return gyro_lnlike + iso_lnlike + lnprior(params)
 
 
-def lnprior(params):
+def lnprior(params, *args):
     """
     lnprior on all parameters.
     """
-    N = int(len(params - 3)/5)  # Number of stars
-    age_prior = sum([np.log(priors.age_prior(np.log10(1e9*np.exp(i))))
-                     for i in params[3+N:3+2*N]])
-    feh_prior = sum([np.log(priors.feh_prior(i)) for i in
-                     params[3+2*N:3+3*N]])
-    distance_prior = sum([np.log(priors.distance_prior(np.exp(i))) for i
-                          in params[3+3*N:3+4*N]])
-    g_prior = priors.lng_prior(params[:3])
+    par_inds = args[-1]
+
+    # if gyro.
+    if par_inds[0] == 0 and par_inds[1] == 1 and par_inds[2] == 2:
+        g_prior = priors.lng_prior(params[:3])
+        age_prior = sum([np.log(priors.age_prior(np.log10(1e9*np.exp(i))))
+                        for i in params[3:]])
+        feh_prior = 0.
+        distance_prior = 0.
+        mAv = True
+
+    # If individual stars
+    else:
+        g_prior = 0.
+        age_prior = np.log(priors.age_prior(np.log10(1e9*np.exp(params[1]))))
+        feh_prior = np.log(priors.feh_prior(params[2]))
+        distance_prior = np.log(priors.distance_prior(np.exp(params[3])))
+
     m = (-20 < params) * (params < 20)  # Broad bounds on all params.
-    Av = params[3+4*N:3+5*N]
-    mAv = (0 <= Av) * (Av < 1)
-    if sum(m) == len(m) and sum(mAv) == len(mAv):
-        return age_prior + feh_prior + distance_prior + g_prior #+ Av_prior
+    mAv = (0 <= params[4]) * (params[4] < 1)  # Prior on A_v
+
+    if sum(m) == len(m) and mAv:
+        return g_prior + age_prior + feh_prior + distance_prior
     else:
         return -np.inf
 
@@ -114,7 +124,7 @@ def lnprob(params, *args):
         args to pass to lnlike, including mods, a list of pre-computed star
         model objects.
     """
-    return lnlike(params, *args) + lnprior(params)
+    return lnlike(params, *args) + lnprior(params, *args)
 
 
[…]

### 2017-05-02 chronometer/chronometer.py
fixing small bug in lnprior

@@ -105,9 +105,9 @@ def lnprior(params, *args):
         age_prior = np.log(priors.age_prior(np.log10(1e9*np.exp(params[1]))))
         feh_prior = np.log(priors.feh_prior(params[2]))
         distance_prior = np.log(priors.distance_prior(np.exp(params[3])))
+        mAv = (0 <= params[4]) * (params[4] < 1)  # Prior on A_v
 
     m = (-20 < params) * (params < 20)  # Broad bounds on all params.
-    mAv = (0 <= params[4]) * (params[4] < 1)  # Prior on A_v
 
     if sum(m) == len(m) and mAv:
         return g_prior + age_prior + feh_prior + distance_prior
@@ -198,7 +198,7 @@ def MH_step(par, lnprob, t, *args, emc=False):
     return par, new_lnprob, accept
 
 
-def gibbs_control(par, lnprob, nsteps_list, niter, t, par_inds_list, args):
+def gibbs_control(par, lnprob, nsteps, niter, t, par_inds_list, args):
     """
     This function tells the metropolis hastings what parameters to sample in.
     params:
@@ -232,7 +232,7 @@ def gibbs_control(par, lnprob, nsteps_list, niter, t, par_inds_list, args):
     n_parameter_sets = len(par_inds_list)
 
     # Final sample array
-    all_samples = np.zeros((sum(nsteps_list) * niter, ndim + nstars))
+    all_samples = np.zeros((nsteps * niter, ndim + nstars))
 
     # Iterate over niter cycles of parameter sets.
     probs = []
@@ -240,8 +240,6 @@ def gibbs_control(par, lnprob, nsteps_list, niter, t, par_inds_list, args):
         print("Gibbs iteration ", i, "of ", niter)
         print(par)
         for k in range(len(par_inds_list)):  # loop over parameter sets.
-            nsteps = nsteps_list[k]
-            print(nsteps, "steps", "parameter_indices = ", par_inds_list[k])
             args[-1] = par_inds_list[k]
             samples, par, pb = MH(par, lnprob, nsteps, t[k],
                                                    *args)
@@ -296,8 +294,8 @@ if __name__ == "__main__":
     start = time.time()  # timeit
 
     # Different nsteps for different parameters. gyro, star 1, 2, 3.
-    nsteps = [1000, 1000, 1000, 1000]
-    niter = 4
+    nsteps = 10000
+    niter = 10
[…]

### 2017-05-02 chronometer/mini_steps_chronometer.py
fixing bugs and testing

@@ -122,8 +122,8 @@ def MH(par, lnprob, nsteps, t, *args):
 
     accept, probs = 0, []
     for i in range(nsteps):
-        par, new_prob, acc = MH_step(par[par_inds], lnprob, t[par_inds],
-                                     *args)
+        par[par_inds], new_prob, acc = MH_step(par, par[par_inds], lnprob,
+                                               t[par_inds], *args)
         accept += acc
         probs.append(new_prob)
         samples[i, :] = par[par_inds]
@@ -131,7 +131,7 @@ def MH(par, lnprob, nsteps, t, *args):
     return samples, par, probs
 
 
-def MH_step(par1, lnprob, t, *args):
+def MH_step(params, par1, lnprob, t, *args):
     """
     A single Metropolis Hastings step.
     if emc = True, the step is an emcee step instead.
@@ -141,7 +141,7 @@ def MH_step(par1, lnprob, t, *args):
     """
     par_ind = args[-1]
     one_par = par1 + np.random.randn(1) * t
-    params = args[-2]*1
+    # params = args[-2]*1
     newp = args[-2]*1
     newp[par_ind] = one_par
     new_lnprob = lnprob(newp, *args)
@@ -157,7 +157,7 @@ def MH_step(par1, lnprob, t, *args):
         else:
             accept = 0
             new_lnprob = lnprob(params, *args)
-    return params, new_lnprob, accept
+    return params[par_ind], new_lnprob, accept
 
 
 def gibbs_control(par, lnprob, nsteps, niter, t, par_inds_list, args):
@@ -199,7 +199,6 @@ def gibbs_control(par, lnprob, nsteps, niter, t, par_inds_list, args):
     probs = []
     for i in range(niter):  # Loop over Gibbs repeats.
         print("Gibbs iteration ", i, "of ", niter)
-        print(par)
         for k in range(len(par_inds_list)):  # loop over parameter sets.
             args[-1] = par_inds_list[k]
             samples, par, pb = MH(par, lnprob, nsteps, t, *args)
@@ -245,12 +244,12 @@ if __name__ == "__main__":
 
     start = time.time()  # timeit
 
-    nsteps, niter = 1000, 5
+    nsteps, niter = 100000, 20
 
     # Construct parameter indices for the different parameter sets.
     par_inds_list = np.arange(len(params))
 
[…]

### 2017-05-02 chronometer/chronometer.py
adding step_chronometer.py

@@ -33,22 +33,20 @@ plt.rcParams.update(plotpar)
 def lnlike(params, *args):
     """
     Probability of age and model parameters given rotation period and colour.
-    parameters:
     Parameters to pass to the models are selected using the par_inds list in
     *args.
-    Whether an "all parameter", "gyro parameter only" or "single star
-    isochrone parameters only" gibbs step is determined according to the what
-    the elements of par_inds look like.
-    This is probably a bad idea.
+    Whether a 'gyro parameter only' or 'single star isochrone parameters only'
+    gibbs step is determined according to the elements of par_inds.
     ----------
     params: (array)
-        The array of log parameters: a, b, n, age.
+        The array of log parameters: a, b, n, masses, ages, etc.
     args = mods, periods, period_errs, bvs, bv_errs, par_inds
-    par_inds: (array)
-        The parameters to vary.
+    mods: (list)
+        A list of starmodel objects, one for each star.
+    par_inds: (array) args[-1]
+        The indices of the parameters to vary.
     """
     mods, period, period_errs, bv, bv_errs, par_inds = args
-    N = len(period)
     gyro_lnlike, iso_lnlike = 0, 0
     if par_inds[0] == 0 and par_inds[1] == 1 and par_inds[2] == 2: # if gyro.
         gyro_lnlike = sum(-.5*((period - gyro_model(params, bv))
@@ -168,7 +166,7 @@ def MH(par, lnprob, nsteps, t, *args):
 
 def MH_step(par, lnprob, t, *args, emc=False):
     """
-    A single Metropolis Hastings step.
+    A single Metropolis step.
     if emc = True, the step is an emcee step instead.
     emcee is run for 10 steps with 64 walkers and the final position is taken
     as the step.
@@ -238,7 +236,7 @@ def gibbs_control(par, lnprob, nsteps, niter, t, par_inds_list, args):
     probs = []
     for i in range(niter):  # Loop over Gibbs repeats.
         print("Gibbs iteration ", i, "of ", niter)
-        print(par)
+        print("Current parameter values = ", par)
         for k in range(len(par_inds_list)):  # loop over parameter sets.
             args[-1] = par_inds_list[k]
             samples, par, pb = MH(par, lnprob, nsteps, t[k],

### 2017-05-02 chronometer/gibbs_chronometer.py

@@ -1,364 +1,0 @@
-"""
-Now use Gibbs sampling to update individual star parameters and global gyro
-parameters.
-"""
-
-import os
-
-import time
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas as pd
-from isochrones import StarModel
-from isochrones.mist import MIST_Isochrone
-
-import corner
-import priors
-from models import gc_model
-
-import teff_bv as tbv
-from utils import replace_nans_with_inits, vk2teff, make_param_dict, \
-    parameter_assignment, pars_and_mods, transform_parameters
-
-plotpar = {'axes.labelsize': 18,
-           'font.size': 10,
-           'legend.fontsize': 18,
-           'xtick.labelsize': 18,
-           'ytick.labelsize': 18,
-           'text.usetex': True}
-plt.rcParams.update(plotpar)
-
-
-def gc_lnlike(params, period, period_errs, bv, bv_errs, all_params=False):
-    """
-    Probability of age and model parameters given rotation period and colour.
-    parameters:
-    ----------
-    params: (array)
-        The array of log parameters: a, b, n, age.
-    period: (tuple)
-        The rotation period and period uncertainty in days
-    bv: (tuple)
-        The B-V colour and colour uncertainty.
-    """
-    pars, ln_ages = transform_parameters(params, "gyro", all_params)
-    model_periods = gc_model(pars, ln_ages, bv)
-    return sum(-.5*((period - model_periods)/period_errs)**2)
-
-
-def iso_lnlike(lnparams, mods, all_params=False):
-    """
-    Some isochronal likelihood function.
-    parameters:
-    ----------
-    params: (array)
-        The array of parameters: log(mass), log(age in Gyr), metallicity,
-        log(distance), extinction.
-    mod: (object)
-        An isochrones.py starmodel object.
-    """
-    p, N = transform_parameters(lnparams, "iso", all_params)
-    if len(mods) > 1:
-        ll = [mods[i].lnlike(p[i::N]) for i in range(len((mods)))]
-        return sum(ll)
-    else:
-        return mods[0].lnlike(p)
-
-
-def gyro_lnprior(params):
-    m = (-10 < params) * (params < 10)  # Broad bounds on all params.
-    if sum(m) == len(m):
-        return priors.lng_prior(params[:3])
-    else:
-        return -np.inf
-
-
-def iso_lnprior(params):
[…]

### 2017-05-02 chronometer/mh_gibbs.py

@@ -1,114 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.mlab as mlab
-import scipy.misc as spm
-import corner
-import emcee
-
-
-def model(par, x):
-    return par[0] + par[1]*x
-
-
-def lnlike(par, x, y, yerr):
-    y_mod = model(par, x)
-    return sum(-.5*((y_mod - y)/yerr)**2)
-
-
-def MH(par, N, t, x, y, yerr):
-    """
-    params:
-    -------
-    par: (list)
-        The parameters.
-    x, y, yerr: (arrays)
-        The data
-    N: (int)
-        Number of samples.
-    t: (float)
-        The std of the proposal distribution.
-    """
-    samples = np.zeros((N, len(par)))
-    for i in range(N):
-        newp = par + np.random.randn(len(par))*t
-        alpha = np.exp(lnlike(newp, x, y, yerr))/np.exp(lnlike(par, x, y,
-                                                               yerr))
-        if alpha > 1:
-            par = newp*1
-        else:
-            u = np.random.uniform(0, 1)
-            if alpha > u:
-                par = newp*1
-        samples[i, :] = par
-    return samples
-
-
-if __name__ == "__main__":
-
-    # Straight line model
-    x = np.arange(0, 10, .1)
-    err = 2.
-    yerr = np.ones_like(x) * err
-    y = .7 + 2.5*x + np.random.randn(len(x))*err
-
-    # Plot the data.
-    plt.clf()
-    plt.errorbar(x, y, yerr=yerr, fmt="k.")
-    plt.savefig("data")
-
-    # First, try brute force random sampling.
-    m = np.random.randn(1000) + 2.5
-    c = np.random.randn(1000) + .5
-    args = np.array([x, y, yerr])
-    Lm = np.array([np.exp(lnlike([.7, m[i]], x, y, yerr)) for i in
-                   range(len(m))])
-    Lc = np.array([np.exp(lnlike([c[i], 2.5], x, y, yerr)) for i in
-                   range(len(m))])
-
-    # Print and plot the results
-    lc, lm = Lc == max(Lc), Lm == max(Lm)
-    print(m[lm], c[lc])
-    plt.clf()
-    plt.plot(m, Lm, "k.")
-    plt.axvline(2.5)
[…]

### 2017-05-02 chronometer/more_stars_chronometer.py

@@ -1,300 +1,0 @@
-# Assembling likelihood functions for Gyrochronology, isochrones, dynamics
-# and asteroseismology.
-
-import os
-
-import time
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas as pd
-import scipy.misc as spm
-from isochrones import StarModel
-from isochrones.mist import MIST_Isochrone
-
-from simple_gibbs import gibbs
-import emcee
-import corner
-import priors
-from models import gc_model
-
-import teff_bv as tbv
-from utils import replace_nans_with_inits, vk2teff, make_param_dict
-
-plotpar = {'axes.labelsize': 18,
-           'font.size': 10,
-           'legend.fontsize': 18,
-           'xtick.labelsize': 18,
-           'ytick.labelsize': 18,
-           'text.usetex': True}
-plt.rcParams.update(plotpar)
-
-
-def gc_lnlike(params, period, period_errs, bv, bv_errs, all_params=False):
-    """
-    Probability of age and model parameters given rotation period and colour.
-    parameters:
-    ----------
-    params: (array)
-        The array of log parameters: a, b, n, age.
-    period: (tuple)
-        The rotation period and period uncertainty in days
-    bv: (tuple)
-        The B-V colour and colour uncertainty.
-    """
-    if all_params:
-        N = int((len(params) - 3)/5)
-        pars = params[:3]
-        ln_ages = params[3+N:3+2*N]
-    else:
-        pars = params[:3]
-        ln_ages = params[3:]
-    model_periods = gc_model(pars, ln_ages, bv)
-    return sum(-.5*((period - model_periods)/period_errs)**2)
-
-
-def iso_lnlike(lnparams, mods, all_params=False):
-    """
-    Some isochronal likelihood function.
-    parameters:
-    ----------
-    params: (array)
-        The array of parameters: log(mass), log(age in Gyr), metallicity,
-        log(distance), extinction.
-    mod: (object)
-        An isochrones.py starmodel object.
-    """
-    if all_params:
-        p = lnparams[3:]*1
-    else:
-        p = lnparams*1
-    N = int(len(p)/5)
-
-    # Transform to linear space
-    # mass, age, feh, distance, Av
[…]

### 2017-05-02 chronometer/multi_star.py

@@ -1,167 +1,0 @@
-# Assembling likelihood functions for Gyrochronology, isochrones, dynamics
-# and asteroseismology.
-
-import os
-
-import time
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas as pd
-import scipy.misc as spm
-from isochrones import StarModel
-from isochrones.mist import MIST_Isochrone
-
-from simple_gibbs import gibbs
-import emcee
-import corner
-
-
-def gc_model(params, bv):
-    """
-    Given a B-V colour and an age predict a rotation period.
-    Returns log(age) in Myr.
-    parameters:
-    ----------
-    params: (array)
-        The array of age and (log) gyro parameters, a, b and n.
-    data: (array)
-        A an array containing colour.
-    """
-    a, b, n, age = np.exp(params)
-    return a*(age*1e3)**n * (bv - .4)**b
-
-
-def gc_lnlike(params, period, bv):
-    """
-    Probability of age and model parameters given rotation period and colour.
-    parameters:
-    ----------
-    params: (array)
-        The array of log parameters: a, b, n, age.
-    period: (tuple)
-        The rotation period and period uncertainty in days
-    bv: (tuple)
-        The B-V colour and colour uncertainty.
-    """
-    model_periods = gc_model(params, bv[0])
-    return sum(-.5*((period[0] - model_periods)/period[1])**2)
-
-
-def iso_lnlike(lnparams, mods):
-    """
-    Some isochronal likelihood function.
-    parameters:
-    ----------
-    params: (array)
-        The array of parameters: log(mass), log(age in Gyr), metallicity,
-        log(distance), extinction.
-    mods: (list)
-        A list of isochrones.py starmodel objects.
-    """
-    params = np.array([np.exp(lnparams[0]), np.log10(1e9*np.exp(lnparams[1])),
-                      lnparams[2], np.exp(lnparams[3]), lnparams[4]])
-    ll = [mod.lnlike(params) for mod in mods]
-    return sum(ll)
-
-
-def lnprior(params):
-    """
-    Some simple, uninformative prior over the parameters.
-    a, b, n, age, mass, feh, distance, Av = params
-    """
-    if -10 < params[0] < 10 and -10 < params[1] < 10 and \
[…]

### 2017-05-02 chronometer/sampling_all_parameters.py

@@ -1,219 +1,0 @@
-"""
-A version where all the parameters are parsed to the lnprob function.
-Doesn't work very well.
-"""
-
-
-import os
-
-import time
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas as pd
-import scipy.misc as spm
-from isochrones import StarModel
-from isochrones.mist import MIST_Isochrone
-
-from simple_gibbs import gibbs
-import emcee
-import corner
-import priors
-from models import gc_model
-
-
-def gc_lnlike(params, period, period_errs, bv, bv_errs):
-    """
-    Probability of age and model parameters given rotation period and colour.
-    parameters:
-    ----------
-    params: (array)
-        The array of log parameters: a, b, n, age.
-    period: (tuple)
-        The rotation period and period uncertainty in days
-    bv: (tuple)
-        The B-V colour and colour uncertainty.
-    """
-    N = int(len(params[3:])/5)  # number of stars
-    ln_age = params[3+N:3+2*N]  # parameter assignment
-    model_periods = gc_model(params[:3], ln_age, bv)
-    return sum(-.5*((period - model_periods)/period_errs)**2)
-
-
-def iso_lnlike(lnparams, mods):
-    """
-    Some isochronal likelihood function.
-    parameters:
-    ----------
-    params: (array)
-        The array of parameters: log(mass), log(age in Gyr), metallicity,
-        log(distance), extinction.
-    mod: (object)
-        An isochrones.py starmodel object.
-    """
-    p = lnparams[3:]*1
-    N = int(len(p)/5)
-
-    # Transform to linear space
-    p[:N] = np.exp(p[:N])
-    p[N:2*N] = np.log10(1e9*np.exp(p[N:2*N]))
-    p[3*N:4*N] = np.exp(p[3*N:4*N])
-
-    ll = [mods[i].lnlike(p[i::N]) for i in range(len((mods)))]
-    return sum(ll)
-
-
-def lnprior(params):
-    N = int(len(params[3:])/5)  # number of stars
-    ln_age = params[3+N:3+2*N]  # parameter assignment
-    ln_mass = params[3+2*N:3+3*N]
-    feh = params[3+3*N:3+4*N]
-    d = params[3+4*N:3+5*N]
-    Av = params[3+5*N:3+6*N]
-    age_prior = sum([np.log(priors.age_prior(np.log10(1e9*np.exp(i))))
[…]

### 2017-05-02 chronometer/simple_gibbs.py

@@ -1,172 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import emcee
-import corner
-
-
-def model(pars, fixed, x, fit):
-    """
-    The Model. A straight line plus a Gaussian.
-    params:
-    ------
-    pars: (tuple, list or string)
-        The parameters to vary.
-    fixed: (tuple, list or string)
-        The fixed parameters.
-    x: (array)
-        The x-array
-    fit: (int)
-        Either 0 or 1. If 0, the line parameters are read as pars and the
-        Gaussian parameters are read as fixed. The reverse is true if 1.
-    returns: (array)
-        The computed y-values of the model.
-    """
-    if fit == 0:
-        a, b = np.exp(pars)
-        mu, sig = np.exp(fixed)
-    elif fit == 1:
-        mu, sig = np.exp(pars)
-        a, b = np.exp(fixed)
-    return a + b*x + np.exp(-(x - mu)**2/(2.*sig**2))
-
-
-def line_lnprob(pars0, pars1, x, y, yerr):
-    """
-    The log-prob of the line model.
-    pars0: (tuple, list or string)
-        The line parameters, pars0 = [a, b]
-    pars1: (tuple, list or string)
-        The Gaussian parameters, pars1 = [mu, sigma]
-    x: (array)
-        The x-array
-    y: (array)
-        The observed y-values
-    yerr: (array)
-        The observed y-uncertainties.
-    returns: (float)
-        The log-probability.
-    """
-    mod = model(pars0, pars1, x, 0)
-    return sum(-((y - mod)**2/(2.*yerr**2))) + line_lnprior(pars0)
-
-
-def line_lnprior(pars0):
-    """
-    The log-prior for the parameters of the line model.
-    """
-    a, b = pars0
-    if -10 < a < 10 and -10 < b < 10:
-        return 0.
-    else:
-        return -np.inf
-
-
-def Gauss_lnprob(pars1, pars0, x, y, yerr):
-    """
-    The log-prob of the Gaussian model.
-    """
-    mod = model(pars1, pars0, x, 1)
-    return sum(-((y - mod)**2/(2.*yerr**2))) + Gauss_lnprior(pars1)
-
-
-def Gauss_lnprior(pars1):
-    """
-    The log-prior for the parameters of the Gaussian model.
-    """
-    mu, sig = pars1
-    if -10 < mu < 10 and -10 < sig < 10:
-        return 0.
-    else:
[…]

### 2017-05-02 chronometer/single_chronometer.py

@@ -1,246 +1,0 @@
-"""
-Assembling likelihood functions for Gyrochronology, isochrones, dynamics
-and asteroseismology.
-"""
-
-import os
-
-import time
-import numpy as np
-import matplotlib.pyplot as plt
-import h5py
-
-import emcee
-import corner
-from isochrones import StarModel
-from isochrones.mist import MIST_Isochrone
-
-import priors
-
-plotpar = {'axes.labelsize': 20,
-           'font.size': 20,
-           'legend.fontsize': 20,
-           'xtick.labelsize': 15,
-           'ytick.labelsize': 15,
-           'text.usetex': True}
-plt.rcParams.update(plotpar)
-
-
-def GC_model(params, bv):
-    """
-    Given a B-V colour and an age predict a rotation period.
-    Returns log(age) in Myr.
-    parameters:
-    ----------
-    params: (array)
-        The array of age and (log) gyro parameters, a, b and n.
-    data: (array)
-        A an array containing colour.
-    """
-    a, b, n, ln_age = params
-    a, b, n = [.7725, .601, .5189]
-    return a*(np.exp(ln_age)*1e3)**n * (bv - .4)**b
-
-
-def gc_lnlike(params, period, bv):
-    """
-    Probability of age and model parameters given rotation period and colour.
-    parameters:
-    ----------
-    params: (array)
-        The array of log parameters: a, b, n, age.
-    period: (tuple)
-        The rotation period and period uncertainty in days
-    bv: (tuple)
-        The B-V colour and colour uncertainty.
-    """
-    model_periods = GC_model(params[:4], bv[0])
-    return -.5*((period[0] - model_periods)/period[1])**2
-
-
-def iso_lnlike(lnparams, mod):
-    """
-    Some isochronal likelihood function.
-    parameters:
-    ----------
-    params: (array)
-        The array of parameters: log(mass), log(age in Gyr), metallicity,
-        log(distance), extinction.
-    mod: (object)
-        An isochrones.py starmodel object.
-    """
-    a, b, n, lnage, lnmass, feh, lnd, Av = lnparams
-    params = np.array([np.exp(lnmass), np.log10(1e9*np.exp(lnage)), feh,
-                       np.exp(lnd), Av])
[…]

### 2017-05-02 chronometer/step_chronometer.py

@@ -1,0 +1,238 @@
+"""
+Now use Gibbs sampling to update individual star parameters and global gyro
+parameters.
+"""
+
+import os
+
+import time
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+from isochrones import StarModel
+from isochrones.mist import MIST_Isochrone
+
+import corner
+import priors
+from models import gc_model, gyro_model
+
+import teff_bv as tbv
+from utils import replace_nans_with_inits, vk2teff, make_param_dict, \
+    parameter_assignment, pars_and_mods, transform_parameters
+
+plotpar = {'axes.labelsize': 18,
+           'font.size': 10,
+           'legend.fontsize': 18,
+           'xtick.labelsize': 18,
+           'ytick.labelsize': 18,
+           'text.usetex': True}
+plt.rcParams.update(plotpar)
+
+
+def lnlike(params, *args):
+    """
+    Probability of age and model parameters given rotation period and colour.
+    parameters:
+    ----------
+    params: (array)
+        The array of log parameters: a, b, n, age.
+    par_inds: (array)
+        The parameters to vary.
+    """
+    mods, period, period_errs, bv, bv_errs, par_inds = args
+    N = len(mods)
+    g_par_inds = np.concatenate((np.arange(3), np.arange(3+N, 3+2*N)))
+    gyro_lnlike =  sum(-.5*((period - gyro_model(params[g_par_inds],
+                                        bv))/period_errs)**2)
+    p = params*1
+    p[3:3+N] = np.exp(p[3:3+N])
+    p[3+N:3+2*N] = np.log10(1e9*np.exp(p[3+N:3+2*N]))
+    p[3+3*N:3+4*N] = np.exp(p[3+3*N:3+4*N])
+    iso_lnlike = sum([mods[i].lnlike(p[3+i::N]) for i in range(len(mods))])
+    return gyro_lnlike + iso_lnlike
+
+
+def lnprior(params):
+    """
+    lnprior on all parameters.
+    """
+    N = int(len(params - 3)/5)  # Number of stars
+    age_prior = sum([np.log(priors.age_prior(np.log10(1e9*np.exp(i))))
+                     for i in params[3+N:3+2*N]])
+    feh_prior = sum([np.log(priors.feh_prior(i)) for i in
+                     params[3+2*N:3+3*N]])
+    distance_prior = sum([np.log(priors.distance_prior(np.exp(i))) for i
+                          in params[3+3*N:3+4*N]])
[…]

### 2017-05-02 chronometer/three_star_example.py

@@ -1,299 +1,0 @@
-# Assembling likelihood functions for Gyrochronology, isochrones, dynamics
-# and asteroseismology.
-
-import os
-
-import time
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas as pd
-import scipy.misc as spm
-from isochrones import StarModel
-from isochrones.mist import MIST_Isochrone
-
-from simple_gibbs import gibbs
-import emcee
-import corner
-import priors
-from models import gc_model
-
-import teff_bv as tbv
-from utils import replace_nans_with_inits, vk2teff, make_param_dict
-
-plotpar = {'axes.labelsize': 18,
-           'font.size': 10,
-           'legend.fontsize': 18,
-           'xtick.labelsize': 18,
-           'ytick.labelsize': 18,
-           'text.usetex': True}
-plt.rcParams.update(plotpar)
-
-
-def gc_lnlike(params, period, period_errs, bv, bv_errs, all_params=False):
-    """
-    Probability of age and model parameters given rotation period and colour.
-    parameters:
-    ----------
-    params: (array)
-        The array of log parameters: a, b, n, age.
-    period: (tuple)
-        The rotation period and period uncertainty in days
-    bv: (tuple)
-        The B-V colour and colour uncertainty.
-    """
-    if all_params:
-        N = int((len(params) - 3)/5)
-        pars = params[:3]
-        ln_ages = params[3+N:3+2*N]
-    else:
-        pars = params[:3]
-        ln_ages = params[3:]
-    model_periods = gc_model(pars, ln_ages, bv)
-    return sum(-.5*((period - model_periods)/period_errs)**2)
-
-
-def iso_lnlike(lnparams, mods, all_params=False):
-    """
-    Some isochronal likelihood function.
-    parameters:
-    ----------
-    params: (array)
-        The array of parameters: log(mass), log(age in Gyr), metallicity,
-        log(distance), extinction.
-    mod: (object)
-        An isochrones.py starmodel object.
-    """
-    if all_params:
-        p = lnparams[3:]*1
-    else:
-        p = lnparams*1
-    N = int(len(p)/5)
-
-    # Transform to linear space
-    # mass, age, feh, distance, Av
[…]

### 2017-05-02 chronometer/working_chronometer.py

@@ -1,298 +1,0 @@
-# Assembling likelihood functions for Gyrochronology, isochrones, dynamics
-# and asteroseismology.
-
-import os
-
-import time
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas as pd
-import scipy.misc as spm
-from isochrones import StarModel
-from isochrones.mist import MIST_Isochrone
-
-from simple_gibbs import gibbs
-import emcee
-import corner
-import priors
-from models import gc_model
-
-import teff_bv as tbv
-from utils import replace_nans_with_inits, vk2teff, make_param_dict
-
-plotpar = {'axes.labelsize': 18,
-           'font.size': 10,
-           'legend.fontsize': 18,
-           'xtick.labelsize': 18,
-           'ytick.labelsize': 18,
-           'text.usetex': True}
-plt.rcParams.update(plotpar)
-
-
-def gc_lnlike(params, period, period_errs, bv, bv_errs, all_params=False):
-    """
-    Probability of age and model parameters given rotation period and colour.
-    parameters:
-    ----------
-    params: (array)
-        The array of log parameters: a, b, n, age.
-    period: (tuple)
-        The rotation period and period uncertainty in days
-    bv: (tuple)
-        The B-V colour and colour uncertainty.
-    """
-    if all_params:
-        N = int((len(params) - 3)/5)
-        pars = params[:3]
-        ln_ages = params[3+N:3+2*N]
-    else:
-        pars = params[:3]
-        ln_ages = params[3:]
-    return sum(-.5*((period - model_periods)/period_errs)**2)
-    model_periods = gc_model(pars, ln_ages, bv)
-
-
-def iso_lnlike(lnparams, mods, all_params=False):
-    """
-    Some isochronal likelihood function.
-    parameters:
-    ----------
-    params: (array)
-        The array of parameters: log(mass), log(age in Gyr), metallicity,
-        log(distance), extinction.
-    mod: (object)
-        An isochrones.py starmodel object.
-    """
-    if all_params:
-        p = lnparams[3:]*1
-    else:
-        p = lnparams*1
-    N = int(len(p)/5)
-
-    # Transform to linear space
-    # mass, age, feh, distance, Av
[…]

### 2017-05-02 chronometer/working_demo.py

@@ -1,309 +1,0 @@
-# Assembling likelihood functions for Gyrochronology, isochrones, dynamics
-# and asteroseismology.
-
-import os
-
-import time
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas as pd
-import scipy.misc as spm
-from isochrones import StarModel
-from isochrones.mist import MIST_Isochrone
-import h5py
-
-from simple_gibbs import gibbs
-import emcee
-import corner
-import priors
-from models import gc_model
-
-import teff_bv as tbv
-from utils import replace_nans_with_inits, vk2teff, make_param_dict
-
-plotpar = {'axes.labelsize': 18,
-           'font.size': 10,
-           'legend.fontsize': 18,
-           'xtick.labelsize': 18,
-           'ytick.labelsize': 18,
-           'text.usetex': True}
-plt.rcParams.update(plotpar)
-
-
-def gc_lnlike(params, period, period_errs, bv, bv_errs, all_params=False):
-    """
-    Probability of age and model parameters given rotation period and colour.
-    parameters:
-    ----------
-    params: (array)
-        The array of log parameters: a, b, n, age.
-    period: (tuple)
-        The rotation period and period uncertainty in days
-    bv: (tuple)
-        The B-V colour and colour uncertainty.
-    """
-    if all_params:
-        N = int((len(params) - 3)/5)
-        pars = params[:3]
-        ln_ages = params[3+N:3+2*N]
-    else:
-        pars = params[:3]
-        ln_ages = params[3:]
-    model_periods = gc_model(pars, ln_ages, bv)
-    return sum(-.5*((period - model_periods)/period_errs)**2)
-
-
-def iso_lnlike(lnparams, mods, all_params=False):
-    """
-    Some isochronal likelihood function.
-    parameters:
-    ----------
-    params: (array)
-        The array of parameters: log(mass), log(age in Gyr), metallicity,
-        log(distance), extinction.
-    mod: (object)
-        An isochrones.py starmodel object.
-    """
-    if all_params:
-        p = lnparams[3:]*1
-    else:
-        p = lnparams*1
-    N = int(len(p)/5)
-
-    # Transform to linear space
[…]

### 2017-05-02 chronometer/working_chronometer_02.05.17.py
adding working_chronometer

@@ -1,0 +1,398 @@
+"""
+Now use Gibbs sampling to update individual star parameters and global gyro
+parameters.
+"""
+
+import os
+
+import time
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+from isochrones import StarModel
+from isochrones.mist import MIST_Isochrone
+import h5py
+
+import corner
+import priors
+from models import gc_model, gyro_model
+import emcee
+
+from utils import replace_nans_with_inits, vk2teff, make_param_dict, \
+    parameter_assignment, pars_and_mods, transform_parameters
+
+plotpar = {'axes.labelsize': 18,
+           'font.size': 10,
+           'legend.fontsize': 18,
+           'xtick.labelsize': 18,
+           'ytick.labelsize': 18,
+           'text.usetex': True}
+plt.rcParams.update(plotpar)
+
+
+def lnlike(params, *args):
+    """
+    Probability of age and model parameters given rotation period and colour.
+    Parameters to pass to the models are selected using the par_inds list in
+    *args.
+    Whether a 'gyro parameter only' or 'single star isochrone parameters only'
+    gibbs step is determined according to the elements of par_inds.
+    ----------
+    params: (array)
+        The array of log parameters: a, b, n, masses, ages, etc.
+    args = mods, periods, period_errs, bvs, bv_errs, par_inds
+    mods: (list)
+        A list of starmodel objects, one for each star.
+    par_inds: (array) args[-1]
+        The indices of the parameters to vary.
+    """
+    mods, period, period_errs, bv, bv_errs, par_inds = args
+    gyro_lnlike, iso_lnlike = 0, 0
+    if par_inds[0] == 0 and par_inds[1] == 1 and par_inds[2] == 2: # if gyro.
+        gyro_lnlike = sum(-.5*((period - gyro_model(params, bv))
+                                /period_errs)**2)
+
+    # If not gyro but single stars
+    else:
+        mod_inds = par_inds[0] - 3
+        ms = mods[mod_inds]
+        p = params*1
+        p[0] = np.exp(p[0])
+        p[1] = np.log10(1e9*np.exp(p[1]))
+        p[3] = np.exp(p[3])
+        iso_lnlike = ms.lnlike(p)
+    return gyro_lnlike + iso_lnlike
+
+
+def emcee_lnprob(params, *args):
+    """
+    The lnprob function used for emcee sampling. Does not break parameters
+    into sets.
+    """
[…]

### 2017-05-02 chronometer/models.py
starting the action-age model

@@ -35,3 +35,16 @@ def gyro_model(params, bv):
     N = len(bv)
     ln_age = params[3:3+N]
     return a*(np.exp(ln_age)*1e3)**n * (bv - .4)**b
+
+
+def action_age(Jz):
+    """
+    Given a vertical action, calculate an age.
+    Vertical action dispersion increases with time.
+    Vertical action is drawn from a Normal distribution with zero mean and
+    dispersion that is a function of time.
+    Calculate the dispersion from the action and convert that into an age
+    using a power law model.
+    This will need to be mass dependent eventually.
+    """
+    return

### 2017-05-03 chronometer/chronometer.py
adding comments

@@ -198,7 +198,10 @@ def MH_step(par, lnprob, t, *args, emc=False):
 
 def gibbs_control(par, lnprob, nsteps, niter, t, par_inds_list, args):
     """
-    This function tells the metropolis hastings what parameters to sample in.
+    This function tells the metropolis hastings what parameters to sample in
+    and assembles the samples into an array. Because the ages are sampled
+    twice, the gyro age parameters are tacked onto the end of this array.
+    I'm not actually sure if this is the correct thing to do...
     params:
     ------
     par: (list)

### 2017-05-03 paper/ms.tex
adding words drunk on the train

--- 2017-03-17 documents/ms.tex
@@
 % Aims
-Stellar age underpins a wide range of astronomical applications, yet
-remains one of the most challenging properties to infer.
+Stellar age underpins a wide range of astronomical applications but
+remains one of the most challenging stellar properties to infer.
 % Method
@@
     wildly inconsistent age estimates.
-In this work we attempt to combine multiple methods into one coherent model
-    and calibrate that model using galactic dynamics --- a well understood
-    physical process.
-We identify a sample of stars with at least two of the following dating
-    methods available: rotation-dating, isochronal-dating,
-    asteroseismic-dating and dynamical-dating.
+In this work we attempt to combine multiple methods into one coherent model.
+For the first time, we combine apparent magnitudes and colour information,
+    typically used to infer an age using stellar evolutionary tracks, with
+    other age proxies: rotation period and galactic kinematics.
+We identify a sample of stars with rotation periods, colours, spectroscopic
+    parameters and kinematics to calibrate this dating method.
 % Results
@@
 The processes behind the formation of the Milky Way and the objects within it
-are some of the most elusive and complicated topics in astronomy today,
-connected by a common theme: stellar ages.
-Ages provide the key to understanding the evolution of all astrophysical
-objects, but age is, unfortunately, one of the most difficult stellar
-properties to measure.
-Different dating methods often produce inconsistent predictions for the age of
-the same star.
+are some of the most elusive and complicated topics in astronomy today.
+These topics are connected by a common theme: stellar ages.
+Stellar ages provide the key to understanding the evolution of all
+astrophysical objects.
+For main sequence (MS) stars however, age is a difficult property to infer.
+This is predominantly because hydrogen burning stars do not change appreciably
+during their time spent on the MS: a star like the Sun will grow in
+luminosity by around a factor of two before becoming a red giant.
+Luminosity is therefore not a very sensitive probe of age.
+As explained further later in this paper: the Sun's rotation period will vary
+by almost an order of magnitude over its MS lifetime, thus rotation
+period is a much more sensitive age proxy than luminosity or temperature.
+
+In addition to the difficulties imposed by the slow timescale for variability
+withing MS stars, different dating methods often produce inconsistent
+predictions for the age of a star.
+For example, an asteroseismic age will not necessarily agree with a isochronal
+or rotational age.
 This is in part because the underlying processes generating the evolution of
@@
 The various available dating methods can be categorised by the underlying
-physical process they trace --- the radial extent of the hydrogen-burning core
-or the evolving magnetic dynamo for example.
-In addition, they can be classified by their level of empiricism, \ie\ the
-number of free parameters that need to tuned when calibrating these models.
-The physics behind orbital heating, for example, is very well understood and
-has only one tunable parameter.
-On the other hand, magnetic activity evolution is poorly understood and the
-dating methods that are based on it are often entirely empirical.
+physical process they trace.
+For example, evolutionary models track the radial extent of the
+hydrogen-burning core and age-rotation relations model the evolving state of
+the internal magnetic dynamo.
+In addition, dating methods can be classified by their level of empiricism,
+\ie\ the number of free parameters that need to tuned when fitting the models
+to the data.
+The physics behind the evolving luminosity and effective temperature of a star
+as a result of core hydrogen burning is, for example, very well understood and
+does not need calibrating; physics determine these models.
+On the other hand, magnetic activity evolution is poorly understood and must
+be calibrated using available data.
 In table \ref{tab:dating_methods} we provide an overview of various dating
@@
 In this paper we focus on four of these dating methods: rotation,
-asteroseismology, isochrones and galactic dynamics.
-A brief overview of each of these methods is provided in below.
+asteroseismology, stellar evolution models and galactic kinematics.
+A brief overview of each of these methods is provided in the following
+sections.
 
@@
     Metcalfe2016}.
-This has been attributed to an evolving magnetic dynamo: as stars reach an
+This has been attributed to an evolving magnetic dynamo: as stars reach a
     critical Rossby number (the ratio of rotation period to the convective
@@
 
-\subsection{Isochrones}
+\subsection{Stellar Evolution models}
+The ongoing physical processes in the core of a star is reflected externally
+by an increase in luminosity and temperature.
+As hydrogen is converted to helium via nuclear fusion in the core, the mass
+fraction, etc, etc.
+Leading to etc, etc.
 
@@
 
+We combine the following dating methods: rotation, stellar evolution models
+and kinematics into one model via self calibration.
+As an example, take the Sun: a star with temperature $5777K$, $ \log(g) =
+4.44$ $[Fe/H] = 0.0$ and $P_{rot} = 26$ days.
+
 In order to establish the level of inconsistency between dating methods, we

### 2017-05-03 chronometer/variance_toy.py
adding variance_typ

@@ -1,0 +1,19 @@
+"""
+Show that variance = x^2
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+true_var = 5
+N = 100000
+
+xs = np.random.randn(N) * true_var**.5
+
+var_xs = xs**2
+
+plt.clf()
+plt.hist(var_xs, 100)
+plt.axvline(true_var, color="r", alpha=.5)
+plt.axvline(np.median(true_var), color="g", alpha=.5, ls="--")
+plt.savefig("var_test")

### 2017-05-08 paper/ms.tex
adding words and equations to methods section

@@
 \usepackage{pdflscape}
+\usepackage{eqnarray}
 
@@
 \newcommand{\feh}{[Fe/H]}
+\newcommand{\Av}{$A_v$}
+\newcommand{\av}{$A_v$}
 \newcommand{\logg}{log \emph{g}}
@@
     typically used to infer an age using stellar evolutionary tracks, with
-    other age proxies: rotation period and galactic kinematics.
+    other age proxies: rotation period and vertical action.
 We identify a sample of stars with rotation periods, colours, spectroscopic
@@
 
-We combine the following dating methods: rotation, stellar evolution models
-and kinematics into one model via self calibration.
+We combine the following dating methods: gyrochronology, isochrone fitting,
+asteroseismology and galactic kinematics into one model.
+In classical isochrone fitting, one might calculate the probability of age
+($A$), mass ($M$), \feh, distance ($D$) and extinction (\av), given a set of
+apparent magnitudes in different filters, \eg\ $M_V$, $M_J$ and $M_K$
+\begin{equation}
+    p(A, M, [Fe/H], D, A_v | M_V, M_J, M_K) \propto p(M_v, M_J, M_K |
+    A, M, [Fe/H], D, A_v)p(A, M, [Fe/H], D, A_v).
+\end{equation}
+Note that there are no free parameters in the model here --- the model is
+entirely governed by physics and does not need to be calibrated further.
+Similarly, in classical gyrochronology, one might calculate the probability of
+age given rotation period and $B-V$ colour,
+\begin{equation}
+    p(A | P_{rot}, B-V) \propto p(P_{rot}, B-V | A)p(A)
+\end{equation}
+In a gyrochronology model, because our understanding of the physics is
+incomplete, there are some free parameters that must be calibrated.
+In the \citet{Barnes2003} gyrochronology parameterisation, these are $a$, $b$
+and $n$, $A = P^{1/n} - (B-V-0.4)^{b/n}$.
+During a calibration exercise, the observables, $A$, $P_{rot}$ and $B-V$ will
+be known and the parameters, $a$, $b$ and $n$ will be inferred,
+\begin{equation}
+    p(a, b, n | A, P_{rot}, B-V) \propto p(A, P_{rot}, B-V | a,
+    b, n) p(a, b, n).
+\end{equation}
+The vertical action-age relation also has free parameters that must be
+calibrated using observations, $A = \alpha (\sigma_{Jz}^2)^\beta$.
+These free parameters, $\alpha$ and $\beta$ can, again be calibrated via
+\begin{equation}
+    p(\alpha, \beta | \{A\}, \{J_z\}) \propto p(\{A\}, \{J_z\} | \alpha,
+    \beta) p(\alpha, \beta).
+\end{equation}
+
+The unique aspect of our analysis is to combine inference over ages with
+calibration of the free parameters in the gyrochronology and vertical
+action-age relations:
+\begin{eqnarray}
+    &&p(\{A\}, \{M\}, \{[Fe/H]\}, \{D\}, \{A_v\}, a, b, n, \alpha, \beta|
+    \{M_V\}, \{M_J\}, \{M_K\}, \{P_{rot}\}, \{B-V\}, \{J_z\}) \\
+    \propto &&p(\{M_v\}, \{M_J\}, \{M_K\}, \{P_{rot}\},
+    \{B-V\}, \{J_z\} | \{A\}, \{M\}, \{[Fe/H]\}, \{D\}, \{A_v\}, a, b, n,
+    \alpha, \beta) \\
+    &&p(A, M, [Fe/H], D, A_v, a, b, n, \alpha, \beta).
+\end{eqnarray}
+We can use conditional independence to break this down further into the
+following:
+\begin{eqnarray}
+    &&p(\{A\}, \{M\}, \{[Fe/H]\}, \{D\}, \{A_v\}, a, b, n, \alpha, \beta|
+    M_V, M_J, M_K,
+    \{P_{rot}\}, \{B-V\}, \{J_z\}) \\
+    \propto &&p(\{M_v\}, \{M_J\}, \{M_K\} | \{A\}, \{M\}, \{[Fe/H]\}, \{D\},
+    \{A_v\})
+    p(A)p(M)p([Fe/H])p(D)p(A_v) \\
+    &&p(\{P_{rot}\}, \{B-V\}, a, b, n | A)p(a)p(b)p(n) \\
+    &&p(\{J_z\}, \alpha, \beta | \{A\}) p(\alpha)p(\beta).
+\end{eqnarray}
+In other words
+\begin{eqnarray}
+    &&p(\mathrm{Ages~\&~other~parameters}|\mathrm{observables}) \\
+    \propto &&\mathcal{L}_{\mathrm{isochronal~age}} \times
+    \mathcal{L}_{\mathrm{gyro~age}} \times
+    \mathcal{L}_{\mathrm{kinematic~age}} \\
+    && \times p(\mathrm{Ages~\&~other~parameters})
+\end{eqnarray}.
+
 As an example, take the Sun: a star with temperature $5777K$, $ \log(g) =
@@
 Kepler mission is provided by the NASA Science Mission directorate.
-\racomment{Add TGAS \& Apogee acknowledgements.}
+\racomment{Add Gaia sprint, TGAS \& Apogee acknowledgements.}
 

### 2017-05-08 chronometer/chronometer.py
trying to run without gyro but returning to original

@@ -49,8 +49,10 @@ def lnlike(params, *args):
     mods, period, period_errs, bv, bv_errs, par_inds = args
     gyro_lnlike, iso_lnlike = 0, 0
     if par_inds[0] == 0 and par_inds[1] == 1 and par_inds[2] == 2: # if gyro.
-        gyro_lnlike = sum(-.5*((period - gyro_model(params, bv))
-                                /period_errs)**2)
+        r = np.isfinite(period)
+        r_period, r_period_errs, r_bv = period[r], period_errs[r], bv[r]
+        gyro_lnlike = sum(-.5*((r_period - gyro_model(params, r_bv))
+                                /r_period_errs)**2)
 
     # If not gyro but single stars
     else:
@@ -286,7 +288,7 @@ if __name__ == "__main__":
 
     # Load the data for the initial parameter array.
     DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
-    d = pd.read_csv(os.path.join(DATA_DIR, "data_file.csv"))
+    d = pd.read_csv(os.path.join(DATA_DIR, "data_file.csv")
 
     # Generate the initial parameter array and the mods objects from the data.
     params, mods = pars_and_mods(DATA_DIR)
@@ -295,7 +297,7 @@ if __name__ == "__main__":
     start = time.time()  # timeit
 
     # Different nsteps for different parameters. gyro, star 1, 2, 3.
-    nsteps = 10000
+    nsteps = 1000
     niter = 10
     N = len(mods)
 
@@ -396,3 +398,8 @@ if __name__ == "__main__":
               np.log(2.5), np.log(2.5)]
     fig = corner.corner(flat, truths=truths, labels=labels)
     fig.savefig(os.path.join(RESULTS_DIR, "demo_corner_gibbs"))
+
+    f = h5py.File(os.path.join(RESULTS_DIR, "samples.h5"), "w")
+    data = f.create_dataset("samples", np.shape(flat))
+    data[:, :] = flat
+    f.close()

### 2017-05-09 paper/ms.tex
editing method section and adding more words

@@
 In classical isochrone fitting, one might calculate the probability of age
-($A$), mass ($M$), \feh, distance ($D$) and extinction (\av), given a set of
-apparent magnitudes in different filters, \eg\ $M_V$, $M_J$ and $M_K$
-\begin{equation}
-    p(A, M, [Fe/H], D, A_v | M_V, M_J, M_K) \propto p(M_v, M_J, M_K |
-    A, M, [Fe/H], D, A_v)p(A, M, [Fe/H], D, A_v).
-\end{equation}
-Note that there are no free parameters in the model here --- the model is
-entirely governed by physics and does not need to be calibrated further.
+$A$, mass $M$, \feh, distance $D$ and extinction \av, given a set of
+apparent magnitudes in different filters, \eg\ $M_V$, $M_J$ and $M_K$.
+This can be written
+\begin{eqnarray}
+    &&p(A, M, [Fe/H], D, A_v | M_V, M_J, M_K) \\ \nonumber
+    \propto &&p(M_v, M_J, M_K | A, M, [Fe/H], D, A_v) \\ \nonumber
+    &&p(A)p(M)p([Fe/H])p(D)p(A_v).
+\end{eqnarray}
+There are no free parameters in the model here --- the model is entirely
+determined by physics and does not need to be calibrated further.
 Similarly, in classical gyrochronology, one might calculate the probability of
@@
 \end{equation}
-In a gyrochronology model, because our understanding of the physics is
-incomplete, there are some free parameters that must be calibrated.
-In the \citet{Barnes2003} gyrochronology parameterisation, these are $a$, $b$
-and $n$, $A = P^{1/n} - (B-V-0.4)^{b/n}$.
+In a gyrochronology model however our understanding of the physics is
+incomplete and some free parameters must be calibrated.
+In the \citet{Barnes2003} gyrochronology parameterisation, $A =
+P_{\mathrm{rot}}^{1/n} - (B-V-0.4)^{b/n}$, these free parameters are $a$, $b$
+and $n$.
 During a calibration exercise, the observables, $A$, $P_{rot}$ and $B-V$ will
@@
 \begin{equation}
-    p(a, b, n | A, P_{rot}, B-V) \propto p(A, P_{rot}, B-V | a,
-    b, n) p(a, b, n).
+    p(a, b, n | \{A\}, \{P_{rot}\}, \{B-V\}) \propto \{p(A\}, \{P_{rot}\},
+    \{B-V\} | a, b, n) p(a)p(b)p(n).
 \end{equation}
-The vertical action-age relation also has free parameters that must be
-calibrated using observations, $A = \alpha (\sigma_{Jz}^2)^\beta$.
-These free parameters, $\alpha$ and $\beta$ can, again be calibrated via
+The vertical action-age relation, $A = \alpha (\sigma_{Jz}^2)^\beta$,
+also has free parameters that must be calibrated using observations.
+These free parameters, $\alpha$ and $\beta$ can again be calibrated via
 \begin{equation}
@@
     &&p(\{A\}, \{M\}, \{[Fe/H]\}, \{D\}, \{A_v\}, a, b, n, \alpha, \beta|
-    \{M_V\}, \{M_J\}, \{M_K\}, \{P_{rot}\}, \{B-V\}, \{J_z\}) \\
+    \{M_V\}, \{M_J\}, \{M_K\}, \{P_{rot}\}, \{B-V\}, \{J_z\}) \\ \nonumber
     \propto &&p(\{M_v\}, \{M_J\}, \{M_K\}, \{P_{rot}\},
     \{B-V\}, \{J_z\} | \{A\}, \{M\}, \{[Fe/H]\}, \{D\}, \{A_v\}, a, b, n,
-    \alpha, \beta) \\
+    \alpha, \beta) \\ \nonumber
     &&p(A, M, [Fe/H], D, A_v, a, b, n, \alpha, \beta).
@@
     M_V, M_J, M_K,
-    \{P_{rot}\}, \{B-V\}, \{J_z\}) \\
+    \{P_{rot}\}, \{B-V\}, \{J_z\}) \\  \nonumber
     \propto &&p(\{M_v\}, \{M_J\}, \{M_K\} | \{A\}, \{M\}, \{[Fe/H]\}, \{D\},
     \{A_v\})
-    p(A)p(M)p([Fe/H])p(D)p(A_v) \\
-    &&p(\{P_{rot}\}, \{B-V\}, a, b, n | A)p(a)p(b)p(n) \\
+    p(A)p(M)p([Fe/H])p(D)p(A_v) \\ \nonumber
+    &&p(\{P_{rot}\}, \{B-V\}, a, b, n | A)p(a)p(b)p(n) \\ \nonumber
     &&p(\{J_z\}, \alpha, \beta | \{A\}) p(\alpha)p(\beta).
@@
 \end{eqnarray}.
+We can multiply the likelihoods together by making the assumption that the
+probabilities are conditionally independent, \ie\ a gyrochronology age depends
+only on rotation period and colour, not proper motion, for example.
+In reality this assumption is not valid because gyrochronology ages do depend
+on \logg, \teff\ and \feh.
+\racomment{How do we get around this!?}
+
+We explore the parameter space via Gibbs sampling.
+Gibbs sampling is a variation of MCMC that makes use of conditional
+independence within the problem to reduce dimensionality.
+It is often used in hierarchical inference problems such as this, where one
+has a set of global parameters ($a$, $b$, $n$, $\alpha$ \& $\beta$) and sets
+of per-object parameters ($\{A\}$, $\{M\}$, $\{[Fe/H]\}$, $\{D\}$, $\{A_v\}$).
+Instead of updating all parameters simultaneously as in Metropolis,
+Metropolis-Hastings, affine invariant \citep{Goodman2010}, nested, Hamiltonian,
+and many other sampling algorithms, in Gibbs sampling, subsets of parameters
+are updated.
+For example, the gyrochronology parameters will be updated, then the kinematic
+parameters, then the isochronal parameters for star 1, then star 2, etc.
+Since the number of parameters is $N_{gyro} + N_{kinematic} + N_{iso} \times
+N_{stars}$, the number of parameters grows rapidly as stars are added.
 

### 2017-05-11 paper/ms.tex
copying across words written for the ADAP

@@
 
+The age-velocity dispersion relations are relatively well studied and based on
+simple physics.
+There is evidence to suggest that stars in the Milky Way form in the thin disc
+of the galaxy with relatively small vertical velocities and angular momenta
+\racomment{add citations}.
+As time passes these stars are scattered via close encounters with other
+stars and interactions with galactic spiral arms.
+The more time passes, the more scattering events, and stars slowly accumulate
+angular momentum in the vertical direction, known as vertical action.
+This results in a slow heating of stellar orbits.
+Older stars can be identified in Gaia DR1 by integrating their orbits in the
+potential of the Milky Way to convert their proper motions, positions and
+parallaxes into vertical actions.
+It is the {\it dispersion} in vertical action that truly traces age, far
+better than vertical action itself.
+However for single stars only the individual's vertical action is available.
+Nonetheless, $J_z^2$ increases over time for any given star, and is a weak
+tracer of age.
+When studying populations of stars, the vertical action dispersion can be used
+to infer their age relatively precisely (to within 2 Gyr
+\racomment{Ruth, check this}).
+However for single stars vertical action can only constrain ages to within 4
+Gyr.
+
 This paper is laid out as follows.
@@
 \label{sec:method}
+
+{\tt chronometer} extends isochrone fitting by simultaneously fitting an
+age-rotation relation and a age-velocity dispersion relation to stars with
+photometric colors, rotation periods and Gaia proper motions.
+It combines the information available from these three dating techniques, thus
+providing more accurate ages than any one of the techniques alone.
+
+The accuracy of ages calculated with {\tt chronometer} still relies on the
+rotational and kinematic age relations being correct.
+As discussed above, the performance of the age-rotation relations for old
+stars has been called into question in the last few years \citep{Angus2015,
+Vansaders2016, Metcalfe2016}.
+A simple straight-line model for rotational age does not reproduce the ages
+predicted by asteroseismology for stars older than the Sun.
+This phenomenon is attributed to a transitioning magnetic dynamo at a critical
+Rossby number, $Ro$, of 2.6, the solar value \citep{Vansaders2016}.
+As rotation periods slow, $Ro$ decreases until it hits the critical value and
+magnetic braking switches off: stars maintain their rotation period from that
+point onward.
+This restricts the applicability of gyrochronology to young, rapidly rotating
+stars.
+This hypothesis is supported by the existing data, however these data are
+sparse~---~the analyses demonstrating the discrepancies were conducted on the
+small number of main sequence, Solar-like oscillators observed by Kepler in
+short cadence mode with detectable rotation periods: a sample size of around
+20 \racomment{Ruth - check this number}.
+% Furthermore, asteroseismology favors stars with low surface gravities since
+% they show the largest amplitudes of oscillation and the majority of stars in
+% this sample were slightly evolved.
+A larger sample of old main sequence stars with precise ages is required to
+confirm and further characterize the Rossby saturation mechanism introduced by
+\citep{Vansaders2016} which is currently supported by the Kepler data.
+However, old main sequence stars are difficult to age-date with any other
+method than asteroseismology.
+We propose use the age gradients in the kinematic properties of stars to
+confirm the Rossby saturation effect.
+
+A simple test will allow us to distinguish between two magnetic braking
+scenarios in MS stars: the continuous scenario in which stars begin magnetic
+braking at zero-age main sequence and continue until main sequence turn off,
+and the transition scenario in which stars stop braking at a critical Rossby
+number.
+This test involves computing rotation-based ages for stars observed by both
+Kepler and Gaia.
+Each star will have ages computed using the continuous scenario and the
+transition scenario.
+The vertical action dispersion will then be calculated for groups of stars
+with similar rotational age, producing two vertical action-age relations.
+The vertical action dispersion will also be calculated for red giants observed
+by APOGEE with ages from the Cannon \citep{Ness2015}, observed by Gaia.
+This red giant vertical action-age relation will serve as our "ground truth"
+since these ages are relatively robust and well calibrated with asteroseismic
+giants.
+We will compare the both the $\sigma_{J_z}$-age relation for the continuous
+scenario and the $\sigma_{J_z}$-age relation for the transition scenario with
+the red giant $\sigma_{J_z}$-age relation.
+This test will reveal whether continuous magnetic braking or transitional
+magnetic braking provides a better description of the data.
+
+This test will allow us to select an appropriate age rotation relation to be
+used in {\tt chronometer}, our new age-dating method for MS stars.
+{\tt chronometer} combines the following dating methods: gyrochronology,
+isochrone fitting, asteroseismology and galactic kinematics into one model by
+simply summing the log-likelihoods of the three different models.
 
@@
 and many other sampling algorithms, in Gibbs sampling, subsets of parameters
-are updated.
+are updated at a time.
 For example, the gyrochronology parameters will be updated, then the kinematic

### 2017-05-14 chronometer/models.py
add action model

@@ -37,14 +37,15 @@ def gyro_model(params, bv):
     return a*(np.exp(ln_age)*1e3)**n * (bv - .4)**b
 
 
-def action_age(Jz):
+def action_age(par, t, Jz, Jz_err):
     """
     Given a vertical action, calculate an age.
     Vertical action dispersion increases with time.
     Vertical action is drawn from a Normal distribution with zero mean and
     dispersion that is a function of time.
-    Calculate the dispersion from the action and convert that into an age
-    using a power law model.
-    This will need to be mass dependent eventually.
     """
-    return
+    if par > 0:
+        return np.sum(-.5*(Jz**2/(par*t + Jz_err**2)) - \
+            .5*np.log(2*np.pi*(par*t + Jz_err**2)))
+    else:
+        return -np.inf

### 2017-05-14 chronometer/test_models.py
adding test_models.py

@@ -1,0 +1,136 @@
+"""
+Test the models in models.py
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+from models import action_age
+import emcee
+
+plotpar = {'axes.labelsize': 18,
+           'font.size': 10,
+           'legend.fontsize': 18,
+           'xtick.labelsize': 18,
+           'ytick.labelsize': 18,
+           'text.usetex': True}
+plt.rcParams.update(plotpar)
+
+
+def jz_lnprior(beta):
+    if 0 < beta < 1e20:
+        return 0.
+    else:
+        return -np.inf
+
+
+def lnprob(par, t, Jz, Jz_err):
+    return action_age(par, t, Jz, Jz_err) + jz_lnprior(par)
+
+
+def test_action_age_model():
+    DATA_DIR = "/Users/ruthangus/granola/granola/data"
+    d = pd.read_csv("ages_and_actions.csv")
+    m = (d.age.values > 0) * (d.age.values < 14)
+    df = d.iloc[m]
+    jz, t = df.Jz.values, df.age.values
+    jz_err = np.zeros(len(jz))
+    print(action_age(1e4, t, jz, np.zeros(len(jz))))
+    args = [t, jz, jz_err]
+    p0 = 1e4
+    init = p0*1
+
+    nwalkers, nsteps, ndim = 64, 1000, 1
+    p0 = [1e-4*np.random.rand(ndim) + p0 for i in range(nwalkers)]
+    sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob, args=args)
+    print("burning in...")
+    pos, _, _ = sampler.run_mcmc(p0, 500)
+    sampler.reset()
+    print("production run...")
+    sampler.run_mcmc(pos, nsteps)
+
+    flat = np.reshape(sampler.chain, (nwalkers*nsteps, ndim))
+
+    plt.clf()
+    plt.hist(flat[:, 0])
+    plt.savefig("hist_test")
+
+    result = np.median(flat[:, 0])
+
+    return init, result
+
+
+def plot_model(t, jz, init, result):
+    xs = np.linspace(min(t), max(t), 100)
+    plt.clf()
+    # plt.loglog()
+    plt.plot(t, jz**2, "k.")
+    plt.plot(xs, init**.5*xs)
[…]

### 2017-05-14 chronometer/chronometer.py
only pass d to pars_and_mods

@@ -260,13 +260,32 @@ def gibbs_control(par, lnprob, nsteps, niter, t, par_inds_list, args):
     return all_samples, lnprobs
 
 
-def estimate_covariance():
+# def estimate_covariance():
+def estimate_covariance(nstars):
     """
     Return the covariance matrix of the emcee samples.
+    If there are more stars than the three that were used to construct this
+    matrix, repeat the last five columns and rows nstar times.
     """
     with h5py.File("emcee_posterior_samples.h5", "r") as f:
         samples = f["samples"][...]
-    return np.cov(samples, rowvar=False)
+    cov = np.cov(samples, rowvar=False)
+    # return cov
+    print(np.shape(cov))
+    n = (np.shape(cov)[0] - 3)/5
+    print(n)
+    nadd = nstars - n
+    print(nadd)
+    star_cov_column = cov[:, -5:]
+    print(np.shape(star_cov_column))
+    for i in range(nadd):
+        newcov = np.vstack((cov, star_cov_column))
+        print(np.shape(newcov))
+        newcov = np.hstack((cov, star_cov_column.T))
+        print(np.shape(newcov))
+    print(np.shape(newcov))
+    assert 0
+    return newcov
 
 
 def find_optimum():
@@ -288,10 +307,11 @@ if __name__ == "__main__":
 
     # Load the data for the initial parameter array.
     DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
-    d = pd.read_csv(os.path.join(DATA_DIR, "data_file.csv")
+    # d = pd.read_csv(os.path.join(DATA_DIR, "data_file.csv")
+    d = pd.read_csv(os.path.join(DATA_DIR, "data.csv"))
 
     # Generate the initial parameter array and the mods objects from the data.
-    params, mods = pars_and_mods(DATA_DIR)
+    params, mods = pars_and_mods(d)
     # params = find_optimum()
 
     start = time.time()  # timeit
@@ -309,7 +329,9 @@ if __name__ == "__main__":
         par_inds_list.append(par_inds[3+i::N])  # Iso stars.
 
     # Create the covariance matrices.
-    t = estimate_covariance()
+    print(len(mods), "len mods")
+    assert 0
+    t = estimate_covariance(N)
     ts = []
     for i, par_ind in enumerate(par_inds_list):
         ti = np.zeros((len(par_ind), len(par_ind)))

### 2017-05-14 chronometer/utils.py

@@ -96,7 +96,7 @@ def transform_parameters(lnparams, indicator, all_params):
         return p, N
 
 
-def pars_and_mods(DATA_DIR):
+def pars_and_mods(d):
     """
     Create the initial parameter array and the mod objects needed for
     isochrones.py from the data file provided.
@@ -109,7 +109,6 @@ def pars_and_mods(DATA_DIR):
     """
     # The initial parameters
     gc = np.array([.7725, .601, .5189])
-    d = pd.read_csv(os.path.join(DATA_DIR, "data_file.csv"))
 
     d = replace_nans_with_inits(d)
     p0 = np.concatenate((gc, np.log(d.mass.values),

### 2017-05-14 chronometer/make_data_file.py
adding kic_tgas.csv

@@ -1,0 +1,34 @@
+"""
+This script is for generating the input data file.
+"""
+
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+
+
+def create_and_save_df(*args):
+    """
+    Create the pandas dataframe, append dictionaries to it and save it as a
+    csv file.
+    input:
+    -----
+    args:
+        Varying numbers of dictionaries containing data to be added to the
+        dataframe.
+    """
+    frames = [i for i in args]
+    df = pd.concat(frames)
+    df.to_csv(os.path.join(DATA_DIR, "data.csv"))
+
+
+if __name__ == "__main__":
+    DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
+    sun = pd.read_csv(os.path.join(DATA_DIR, "sun.csv"))
+    NGC2019 = pd.read_csv(os.path.join(DATA_DIR, "NGC2019.csv"))
+    tgas = pd.read_csv(os.path.join(DATA_DIR, "kic_tgas.csv"))
+    tgas = tgas.iloc[:2, :]
+
+    # create_and_save_df(sun, NGC2019)
+    create_and_save_df(sun, NGC2019, tgas)

### 2017-05-23 chronometer/actions.py
adding actions and calculate_actions.py

@@ -1,0 +1,142 @@
+# Calculate posterior samples of actions.
+# Code from Wilma Trick's github repo:
+# https://github.com/wilmatrick/GaiaSprint/blob/master/Action_Galpy_Tutorial
+# .ipynb
+
+import numpy as np
+
+from galpy.util import bovy_coords
+from galpy.potential import MWPotential2014 as pot
+from galpy.actionAngle import actionAngleStaeckel
+
+
+def action(ra_deg, dec_deg, d_kpc, pm_ra_masyr, pm_dec_masyr, v_los_kms):
+    """
+    Use Wilma Trick's code to calculate vertical action from
+    params:
+    ------
+    ra_deg: (float)
+        RA in degrees.
+    dec_deg: (float)
+        declination in degrees.
+    d_kpc: (float)
+        Distance in kpc,
+    pm_ra_masyr: (float)
+        RA proper motion in milliarcseconds/year
+    pm_dec_masyr: (float)
+        declination proper motion in milliarcseconds/year
+    v_los_kms: (float)
+        Line of sight velocity in km/s.
+    """
+    ra_rad = ra_deg * (np.pi / 180.)  # RA [rad]
+    dec_rad = dec_deg * (np.pi / 180.)  # dec [rad]
+
+    # Galactocentric position of the Sun:
+    X_gc_sun_kpc = 8.  # [kpc]
+    Z_gc_sun_kpc = 0.025  # [kpc]
+
+    # Galactocentric velocity of the Sun:
+    vX_gc_sun_kms = -9.58  # = -U              [kms]
+    vY_gc_sun_kms = 10.52 + 220.  # = V+v_circ(R_Sun) [kms]
+    vZ_gc_sun_kms = 7.01  # = W               [kms]
+
+    # a. convert spatial coordinates (ra,dec,d) to (R,z,phi)
+
+    # (ra,dec) --> Galactic coordinates (l,b):
+    lb = bovy_coords.radec_to_lb(ra_rad, dec_rad, degree=False, epoch=2000.0)
+    # l_rad = lb[:, 0]
+    # b_rad = lb[:, 1]
+    l_rad = lb[0]
+    b_rad = lb[1]
+
+    # (l,b,d) --> Galactocentric cartesian coordinates (x,y,z):
[…]

### 2017-05-23 chronometer/calculate_actions.py

@@ -1,0 +1,86 @@
+"""
+Take the parallax and proper motions of a TGAS star and calculate Jz.
+Append Jz to a csv file.
+"""
+
+import numpy as np
+import pandas as pd
+from actions import calc_actions
+import os
+
+
+def load_pd_file(fn):
+    df = pd.read_csv(fn)
+    return df
+
+
+def zero_fill(array):
+    """
+    Replace nans with zeros
+    """
+    m = np.isfinite(array)
+    array[~m] = np.zeros(len(array[~m]))
+    return array
+
+
+def convert_coords_to_actions(df):
+    """
+    Take an existing pandas dataframe containing plx and pm info.
+    Convert data to actions.
+    """
+
+    # Load info from dataframe and fill with zeros where unavailable.
+    ra_deg, ra_deg_err = zero_fill(df.tgas_ra), zero_fill(df.tgas_ra_error)
+    dec_deg, dec_deg_err = zero_fill(df.tgas_dec), zero_fill(df.tgas_dec_error)
+    ra_dec_corr = zero_fill(df.tgas_ra_dec_corr)
+    d_kpc = 1./df.tgas_parallax
+    d_kpc_err = df.tgas_parallax_error/df.tgas_parallax * d_kpc
+    d_kpc, d_kpc_err = zero_fill(d_kpc), zero_fill(d_kpc_err)
+    pm_ra_masyr = zero_fill(df.tgas_pmra)
+    pm_ra_masyr_err = zero_fill(df.tgas_pmra_error)
+    pm_dec_masyr = zero_fill(df.tgas_pmdec)
+    pm_dec_masyr_err = zero_fill(df.tgas_pmdec_error)
+    # FIXME: add correlations
+
+    # Look for RVs. If none exist, use zeros for now.
+    try:
+        v_los_kms = df.rv
+        v_los_kms_err = df.rv_err
+    except:
+        v_los_kms = np.zeros(len(ra_deg))
+        v_los_kms_err = np.zeros(len(ra_deg))
+
+
+    # Calculate actions and errors.
+    print(np.shape(ra_deg), np.shape(dec_deg), np.shape(d_kpc),
+          np.shape(pm_ra_masyr), np.shape(pm_dec_masyr), np.shape(v_los_kms))
+    print(ra_deg[0], dec_deg[0], d_kpc[0], pm_ra_masyr[0], pm_dec_masyr[0],
+          v_los_kms[0])
[…]

### 2017-05-23 chronometer/action_age_evolution.py
adding all calculated values to pd.dataframe

@@ -62,6 +62,7 @@ if __name__ == "__main__":
     plt.plot(ages, dispersions)
     plt.xlabel("log(Time)")
     plt.ylabel("dispersion")
+    plt.subplots_adjust(bottom=.15)
     plt.savefig("age_dispersion")
 
     assert 0

### 2017-05-23 chronometer/calculate_actions.py

@@ -5,7 +5,7 @@ Append Jz to a csv file.
 
 import numpy as np
 import pandas as pd
-from actions import calc_actions
+from actions import action
 import os
 
 
@@ -56,31 +56,37 @@ def convert_coords_to_actions(df):
           np.shape(pm_ra_masyr), np.shape(pm_dec_masyr), np.shape(v_los_kms))
     print(ra_deg[0], dec_deg[0], d_kpc[0], pm_ra_masyr[0], pm_dec_masyr[0],
           v_los_kms[0])
-    jr, lz, jz = calc_actions(ra_deg[0], dec_deg[0], d_kpc[0], pm_ra_masyr[0],
-                              pm_dec_masyr[0], v_los_kms[0])
+    R_kpc, phi_rad, z_kpc, vR_kms, vT_kms, vz_kms, jr, lz, jz = \
+        [np.zeros(len(ra_deg)) for i in range(9)]
     jr_err, lz_err, jz_err = [np.zeros(len(jr)) for i in range(3)]
+    for i in range(len(ra_deg)):
+        R_kpc[i], phi_rad[i], z_kpc[i], vR_kms[i], vT_kms[i], vz_kms[i], \
+            jr[i], lz[i], jz[i] = action(ra_deg[i], dec_deg[i], d_kpc[i],
+                                         pm_ra_masyr[i], pm_dec_masyr[i],
+                                         v_los_kms[i])
     print(jz)
     assert 0
-    return jr, jr_err, lz, lz_err, jz, jz_err
+    return np.array([R_kpc, R_kpc_err, phi_rad, phi_rad_err, z_kpc, z_kpc_err,
+                     vR_kms, vR_kms_err, vT_kms, vT_kms_err, vz_kms,
+                     vz_kms_err, jr, jr_err, lz, lz_err, jz, jz_err])
 
 
-def save_pd_file(df, fn, jr, jr_err, lz, lz_err, jz, jz_err):
+def save_pd_file(df, fn, array):
     """
     Add the actions to the df and save it under a different name.
     """
-    df["Jr"] = jr
-    df["Jr_err"] = jr_err
-    df["Lz"] = lz
-    df["Lz_err"] = lz_err
-    df["Jz"] = jz
-    df["Jz_err"] = jz_err
+    df["R_kpc"], df["R_kpc_err"], df["phi_rad"], df["phi_rad_err"], \
+        df["z_kpc"], df["z_kpc_err"], df["vR_kms"], df["vR_kms_err"], \
[…]

### 2017-05-23 chronometer/chronometer.py

@@ -309,6 +309,8 @@ if __name__ == "__main__":
     DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
     # d = pd.read_csv(os.path.join(DATA_DIR, "data_file.csv")
     d = pd.read_csv(os.path.join(DATA_DIR, "data.csv"))
+    print(d.keys())
+    assert 0
 
     # Generate the initial parameter array and the mods objects from the data.
     params, mods = pars_and_mods(d)

### 2017-05-23 chronometer/make_data_file.py

@@ -19,7 +19,7 @@ def create_and_save_df(*args):
         dataframe.
     """
     frames = [i for i in args]
-    df = pd.concat(frames)
+    df = pd.concat(frames, axis=0, how="outer")
     df.to_csv(os.path.join(DATA_DIR, "data.csv"))
 
 

### 2017-05-23 chronometer/mcquillan_granola.py

@@ -58,7 +58,16 @@ if __name__ == "__main__":
     m = (d.age.values > 0) * (d.age.values < 14)
     df = d.iloc[m]
 
-    ages, dispersions, Ns = calc_dispersion(df.age.values, df.Jz.values, 8)
+    plt.clf()
+    plt.plot(np.log(df.Jz.values**2), np.log(df.age.values), "k.")
+    xs = np.linspace(min(df.Jz.values), max(df.Jz.values), 100)
+    ys = .2*xs - (.9/.2)
+    # plt.plot(xs, ys)
+    plt.ylabel("Age")
+    plt.xlabel("Jz2")
+    plt.savefig("age_vs_Jz2")
+
+    ages, dispersions, Ns = calc_bin_dispersion(df.age.values, df.Jz.values, 8)
     d_err = dispersions / (2 * Ns - 2)**.5
 
     plt.clf()
@@ -70,7 +79,7 @@ if __name__ == "__main__":
     plt.savefig("linear_age_dispersion.pdf")
 
     m = np.log(df.age.values) > - 1
-    lnages, dispersions, Ns = calc_dispersion(np.log(df.age.values[m]),
+    lnages, dispersions, Ns = calc_bin_dispersion(np.log(df.age.values[m]),
                                       df.Jz.values[m], 8)
     d_err = dispersions / (2 * Ns - 2)**.5
 
@@ -81,6 +90,9 @@ if __name__ == "__main__":
     plt.xlabel("$\ln(\mathrm{Age,~Gyr})$")
     plt.ylabel("$\sigma J_z~(\mathrm{Kpc~kms}^{-1})$")
     plt.xlim(-1, 2.6)
+    xs = np.linspace(min(lnages), max(lnages), 100)
+    ys = .2*xs + .9
+    plt.plot(xs, ys)
     plt.savefig("log_age_dispersion.pdf")
 
     m = np.log(df.age.values) > - 1

### 2017-05-23 chronometer/utils.py

@@ -27,10 +27,8 @@ def replace_nans_with_inits(data):
     data.Av.values[~np.isfinite(data.Av.values)] \
         = np.zeros(len(data.Av.values[~np.isfinite(data.Av.values)]))
     data.parallax.values[~np.isfinite(data.parallax.values)] \
-        = np.zeros(len(data.parallax.values[~np.isfinite(data.parallax.values)]))
-    # data.parallax.values[~np.isfinite(data.parallax.values)] \
-    #     = [None] * \
-    #     (len(data.parallax.values[~np.isfinite(data.parallax.values)]))
+        = np.zeros(len(data.parallax.values[~np.isfinite(
+            data.parallax.values)]))
     return data
 
 

### 2017-05-23 chronometer/actions.py
fixing array error

@@ -123,13 +123,21 @@ def calc_actions(R_kpc, phi_rad, z_kpc, vR_kms, vT_kms, vz_kms):
     _REFR0 = 8.  # [kpc]  --> galpy length unit
     _REFV0 = 220.  # [km/s] --> galpy velocity unit
 
-    R = np.array([R_kpc, 8.1]) / _REFR0  # Galactocentric radius
-    vR = np.array([vR_kms, 0.]) / _REFV0  # radial velocity
-    phi = np.array([phi_rad, 0.])  # Galactocentric azimuth angle
+    # R = np.array([R_kpc, 8.1]) / _REFR0  # Galactocentric radius
+    # vR = np.array([vR_kms, 0.]) / _REFV0  # radial velocity
+    # phi = np.array([phi_rad, 0.])  # Galactocentric azimuth angle
+    # # (not needed for actions in axisymmetric potential)
+    # vT = np.array([vT_kms, 230.]) / _REFV0  # tangential velocity
+    # z = np.array([z_kpc, 0.1]) / _REFR0  # height above plane
+    # vz = np.array([vz_kms, 0.]) / _REFV0  # vertical velocity
+
+    R = R_kpc / _REFR0  # Galactocentric radius
+    vR = vR_kms / _REFV0  # radial velocity
+    phi = phi_rad  # Galactocentric azimuth angle
     # (not needed for actions in axisymmetric potential)
-    vT = np.array([vT_kms, 230.]) / _REFV0  # tangential velocity
-    z = np.array([z_kpc, 0.1]) / _REFR0  # height above plane
-    vz = np.array([vz_kms, 0.]) / _REFV0  # vertical velocity
+    vT = vT_kms / _REFV0  # tangential velocity
+    z = z_kpc / _REFR0  # height above plane
+    vz = vz_kms / _REFV0  # vertical velocity
 
     # delta = focal length of confocal coordinate system
     # Use C code (for speed)

### 2017-05-23 chronometer/calculate_actions.py

@@ -58,17 +58,22 @@ def convert_coords_to_actions(df):
           v_los_kms[0])
     R_kpc, phi_rad, z_kpc, vR_kms, vT_kms, vz_kms, jr, lz, jz = \
         [np.zeros(len(ra_deg)) for i in range(9)]
-    jr_err, lz_err, jz_err = [np.zeros(len(jr)) for i in range(3)]
+    jr_err, lz_err, jz_err, R_kpc_err, phi_rad_err, z_kpc_err, vR_kms_err, \
+        vT_kms_err, vz_kms_err = [np.zeros(len(jr)) for i in range(9)]
     for i in range(len(ra_deg)):
         R_kpc[i], phi_rad[i], z_kpc[i], vR_kms[i], vT_kms[i], vz_kms[i], \
             jr[i], lz[i], jz[i] = action(ra_deg[i], dec_deg[i], d_kpc[i],
                                          pm_ra_masyr[i], pm_dec_masyr[i],
                                          v_los_kms[i])
-    print(jz)
-    assert 0
-    return np.array([R_kpc, R_kpc_err, phi_rad, phi_rad_err, z_kpc, z_kpc_err,
-                     vR_kms, vR_kms_err, vT_kms, vT_kms_err, vz_kms,
-                     vz_kms_err, jr, jr_err, lz, lz_err, jz, jz_err])
+    m = (ra_deg == 0.) * (dec_deg == 0.) * (pm_ra_masyr == 0.) * \
+        (pm_dec_masyr == 0.)
+    action_array = np.array([R_kpc, R_kpc_err, phi_rad, phi_rad_err, z_kpc,
+                             z_kpc_err, vR_kms, vR_kms_err, vT_kms,
+                             vT_kms_err, vz_kms, vz_kms_err, jr, jr_err, lz,
+                             lz_err, jz, jz_err])
+    for i, arr in enumerate(action_array):  # Mute non-values
+        action_array[i][m] = np.zeros(len(action_array[i][m]))
+    return action_array
 
 
 def save_pd_file(df, fn, array):
@@ -80,6 +85,7 @@ def save_pd_file(df, fn, array):
         df["vT_kms"], df["vT_kms_err"], df["vz_kms"], df["vz_kms_err"], \
         df["Jr"], df["Jr_err"], df["Lz"], df["Lz_err"], df["Jz"], \
         df["Jz_err"] = array
+    print(df["R_kpc"])
     df.to_csv(fn)
 
 
[…]

### 2017-05-23 chronometer/calculate_actions.py
inserting NaNs instead of zeros

@@ -72,7 +72,7 @@ def convert_coords_to_actions(df):
                              vT_kms_err, vz_kms, vz_kms_err, jr, jr_err, lz,
                              lz_err, jz, jz_err])
     for i, arr in enumerate(action_array):  # Mute non-values
-        action_array[i][m] = np.zeros(len(action_array[i][m]))
+        action_array[i][m] = np.ones(len(action_array[i][m])) * np.nan
     return action_array
 
 

### 2017-05-23 chronometer/chronometer.py
adding G to df

@@ -308,27 +308,33 @@ if __name__ == "__main__":
     # Load the data for the initial parameter array.
     DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
     # d = pd.read_csv(os.path.join(DATA_DIR, "data_file.csv")
-    d = pd.read_csv(os.path.join(DATA_DIR, "data.csv"))
-    print(d.keys())
-    assert 0
+    d = pd.read_csv(os.path.join(DATA_DIR, "action_data.csv"))
 
-    # Generate the initial parameter array and the mods objects from the data.
-    params, mods = pars_and_mods(d)
+    # Generate the initial parameter array and the mods objects from the data
+    # The initial parameters
+    global_params = np.array([.7725, .601, .5189, 350.])  # a, b, n, beta
+
+    params, mods = pars_and_mods(d, global_params)
     # params = find_optimum()
 
     start = time.time()  # timeit
 
-    # Different nsteps for different parameters. gyro, star 1, 2, 3.
+    # Set nsteps and niter.
     nsteps = 1000
     niter = 10
-    N = len(mods)
+    N = len(mods)  # number of stars
+    nglob, nind = 4, 5  # number of global pars and number of per-star pars.
+    ngyro = 3  # number of gyro parameters.
 
     # Construct parameter indices for the different parameter sets.
     par_inds = np.arange(len(params))  # All
-    gyro_par_inds = np.concatenate((par_inds[:3], par_inds[3+N:3+2*N])) # Gyro
+    gyro_par_inds = np.concatenate((par_inds[:ngyro],
+                                    par_inds[nglob+N:nglob+2*N])) # Gyro pars
     par_inds_list = [gyro_par_inds]
     for i in range(N):
-        par_inds_list.append(par_inds[3+i::N])  # Iso stars.
+        par_inds_list.append(par_inds[nglob+i::N])  # Iso stars.
+    print(par_inds_list)
+    assert 0
 
     # Create the covariance matrices.
     print(len(mods), "len mods")

### 2017-05-23 paper/ms.tex
trying to run with emcee, first steps

@@
 In this work we attempt to combine multiple methods into one coherent model.
-For the first time, we combine apparent magnitudes and colour information,
-    typically used to infer an age using stellar evolutionary tracks, with
-    other age proxies: rotation period and vertical action.
+We combine apparent magnitudes and colour indices, typically used to infer an
+    age using stellar evolutionary tracks alone, with other age proxies:
+    rotation period and vertical action.
 We identify a sample of stars with rotation periods, colours, spectroscopic
@@
 In addition to the difficulties imposed by the slow timescale for variability
-withing MS stars, different dating methods often produce inconsistent
+within MS stars, different dating methods often produce inconsistent
 predictions for the age of a star.

### 2017-05-23 chronometer/chronometer.py

@@ -71,17 +71,25 @@ def emcee_lnprob(params, *args):
     The lnprob function used for emcee sampling. Does not break parameters
     into sets.
     """
-    mods, period, period_errs, bv, bv_errs, _ = args
-    g_par_inds = np.concatenate((par_inds[:3], par_inds[3+N:3+2*N]))
-    gyro_lnlike =  sum(-.5*((period - gyro_model(params[g_par_inds], bv))
+    ngyro, nglob, N, mods, period, period_errs, bv, bv_errs, Jz, Jz_err, \
+        par_inds = args
+    age_par_inds = par_inds[nglob+N:nglob+2*N]
+    g_par_inds = np.concatenate((par_inds[:ngyro],
+                                 par_inds[nglob+N:nglob+2*N]))
+    print(params, "params")
+    print(g_par_inds, "g_par_inds")
+    print(params[g_par_inds])
+    assert 0
+    gyro_lnlike = sum(-.5*((period - gyro_model(params[g_par_inds], bv))
                             /period_errs)**2)
+    kin_lnlike = action_age(params[ngyro], params[age_par_inds], Jz, Jz_err)
     p = params*1
-    p[3:3+N] = np.exp(p[3:3+N])
-    p[3+N:3+2*N] = np.log10(1e9*np.exp(p[3+N:3+2*N]))
-    p[3+3*N:3+4*N] = np.exp(p[3+3*N:3+4*N])
-    iso_lnlike = sum([mods[i].lnlike(p[3+i::N]) for i in
-                        range(len(mods))])
-    return gyro_lnlike + iso_lnlike + lnprior(params)
+    p[nglob:nglob+N] = np.exp(p[nglob:nglob+N])
+    p[nglob+N:nglob+2*N] = np.log10(1e9*np.exp(p[nglob+N:nglob+2*N]))
+    p[nglob+nglob*N:nglob+4*N] = np.exp(p[nglob+nglob*N:nglob+4*N])
+    iso_lnlike = sum([mods[i].lnlike(p[nglob+i::N]) for i in
+                      range(len(mods))])
+    return gyro_lnlike + iso_lnlike + kin_lnlike + lnprior(params, *args)
 
 
 def lnprior(params, *args):
@@ -98,6 +106,15 @@ def lnprior(params, *args):
         feh_prior = 0.
         distance_prior = 0.
         mAv = True
+
+    # if kinematics
+    if par_inds[0] == 3:
[…]

### 2017-05-23 chronometer/make_data_file.py

@@ -19,7 +19,7 @@ def create_and_save_df(*args):
         dataframe.
     """
     frames = [i for i in args]
-    df = pd.concat(frames, axis=0, how="outer")
+    df = pd.concat(frames, axis=0)
     df.to_csv(os.path.join(DATA_DIR, "data.csv"))
 
 

### 2017-05-23 chronometer/utils.py

@@ -56,16 +56,18 @@ def make_param_dict(d, i):
     i: (int)
         The index of the star.
     """
-    param_dict = {"V": (d.v.values[i], d.v_err.values[i]),
-                  "Ks": (d.ks.values[i], d.ks_err.values[i]),
-                  "J": (d.j.values[i], d.j_err.values[i]),
+    param_dict = {"J": (d.j.values[i], d.j_err.values[i]),
                   "H": (d.h.values[i], d.h_err.values[i]),
                   "K": (d.k.values[i], d.k_err.values[i]),
+                  "G": (d.tgas_phot_g_mean_mag.values[i],
+                        d.tgas_phot_g_mean_flux_error.values[i]/
+                        d.tgas_phot_g_mean_flux.values[i] *
+                        d.tgas_phot_g_mean_mag.values[i]),
                   "Teff": (d.Teff.values[i], d.Teff_err.values[i]),
                   "logg": (d.logg.values[i], d.logg_err.values[i]),
                   "feh": (d.feh.values[i], d.feh_err.values[i]),
                   "parallax": (d.parallax.values[i],
-                               d.parallax_err.values[i])}
+                               d.parallax_err.values[i])}  # FIXME: add more filters
     return param_dict
 
 
@@ -94,7 +96,7 @@ def transform_parameters(lnparams, indicator, all_params):
         return p, N
 
 
-def pars_and_mods(d):
+def pars_and_mods(d, global_params):
     """
     Create the initial parameter array and the mod objects needed for
     isochrones.py from the data file provided.
@@ -105,11 +107,9 @@ def pars_and_mods(d):
     mods: (list)
         A list of starmodel objects.
     """
-    # The initial parameters
-    gc = np.array([.7725, .601, .5189])
 
     d = replace_nans_with_inits(d)
-    p0 = np.concatenate((gc, np.log(d.mass.values),
+    p0 = np.concatenate((global_params, np.log(d.mass.values),
                          np.log(d.age.values), d.feh.values,
                          np.log(1./d.parallax.values*1e3),
                          d.Av.values))

### 2017-05-24 chronometer/utils.py
adding default value for age

@@ -8,6 +8,7 @@ import pandas as pd
 from isochrones import StarModel
 from isochrones.mist import MIST_Isochrone
 
+
 def replace_nans_with_inits(data):
     """
     Turn the data into reasonable initial values.
@@ -20,6 +21,7 @@ def replace_nans_with_inits(data):
     If no initial feh is provided, initialise with solar feh.
     If no initial Av is provided, initialise with 0.
     """
+
     data.mass.values[~np.isfinite(data.mass.values)] \
         = np.ones(len(data.mass.values[~np.isfinite(data.mass.values)]))
     data.feh.values[~np.isfinite(data.feh.values)] \
@@ -29,6 +31,9 @@ def replace_nans_with_inits(data):
     data.parallax.values[~np.isfinite(data.parallax.values)] \
         = np.zeros(len(data.parallax.values[~np.isfinite(
             data.parallax.values)]))
+    data.age.values[~np.isfinite(data.age.values)] \
+        = np.ones(len(data.age.values[~np.isfinite(data.age.values)])) * \
+        np.log(2)
     return data
 
 

### 2017-05-24 chronometer/make_data_file.py
filling in missing B-Vs

@@ -6,9 +6,11 @@ import os
 import numpy as np
 import matplotlib.pyplot as plt
 import pandas as pd
+from calculate_actions import convert_coords_to_actions, save_pd_file
+import teff_bv as tbv
 
 
-def create_and_save_df(*args):
+def create_and_save_df(fn, *args):
     """
     Create the pandas dataframe, append dictionaries to it and save it as a
     csv file.
@@ -20,7 +22,7 @@ def create_and_save_df(*args):
     """
     frames = [i for i in args]
     df = pd.concat(frames, axis=0)
-    df.to_csv(os.path.join(DATA_DIR, "data.csv"))
+    df.to_csv(os.path.join(DATA_DIR, fn))
 
 
 if __name__ == "__main__":
@@ -31,4 +33,14 @@ if __name__ == "__main__":
     tgas = tgas.iloc[:2, :]
 
     # create_and_save_df(sun, NGC2019)
-    create_and_save_df(sun, NGC2019, tgas)
+    fn = "data.csv"
+    create_and_save_df(fn, sun, NGC2019, tgas)
+    df = pd.read_csv(os.path.join(DATA_DIR, fn))
+
+    # Calculate B-Vs
+    bv = tbv.teff2bv(df.teff.values, df.logg.values, df.feh.values)
+    df["bv"] = bv
+
+    # Calculate actions
+    action_array = convert_coords_to_actions(df)
+    save_pd_file(df, os.path.join(DATA_DIR, "action_data.csv"), action_array)

### 2017-05-24 chronometer/utils.py
combining columns

@@ -31,6 +31,9 @@ def replace_nans_with_inits(data):
     data.parallax.values[~np.isfinite(data.parallax.values)] \
         = np.zeros(len(data.parallax.values[~np.isfinite(
             data.parallax.values)]))
+    data.tgas_parallax.values[~np.isfinite(data.tgas_parallax.values)] \
+        = np.zeros(len(data.tgas_parallax.values[~np.isfinite(
+            data.tgas_parallax.values)]))
     data.age.values[~np.isfinite(data.age.values)] \
         = np.ones(len(data.age.values[~np.isfinite(data.age.values)])) * \
         np.log(2)
@@ -101,10 +104,28 @@ def transform_parameters(lnparams, indicator, all_params):
         return p, N
 
 
+def combine_columns(d, col1, col2):
+    """
+    Where there are two columns that need to be combined. Return the combined
+    columns, col1 and col2.
+    d: the pd dataframe.
+    col1: (array)
+        The first column as an array
+    col2: (array)
+        The second ".
+    """
+    a = np.ones(len(col1)) * np.nan
+    m1, m2 = np.isfinite(col1), np.isfinite(col2)
+    a[m1], a[m2] = col1[m1], col2[m2]
+    return a
+
+
 def pars_and_mods(d, global_params):
     """
     Create the initial parameter array and the mod objects needed for
     isochrones.py from the data file provided.
+    Also, perform some operations on the data: merge multiple parallax columns
+    and covert teffs into B-V.
     returns:
     -------
     p0: (array)

### 2017-05-24 chronometer/utils.py
no longer combining columns

@@ -104,22 +104,6 @@ def transform_parameters(lnparams, indicator, all_params):
         return p, N
 
 
-def combine_columns(d, col1, col2):
-    """
-    Where there are two columns that need to be combined. Return the combined
-    columns, col1 and col2.
-    d: the pd dataframe.
-    col1: (array)
-        The first column as an array
-    col2: (array)
-        The second ".
-    """
-    a = np.ones(len(col1)) * np.nan
-    m1, m2 = np.isfinite(col1), np.isfinite(col2)
-    a[m1], a[m2] = col1[m1], col2[m2]
-    return a
-
-
 def pars_and_mods(d, global_params):
     """
     Create the initial parameter array and the mod objects needed for

### 2017-05-24 chronometer/calculate_actions.py
modding docstring

@@ -1,6 +1,7 @@
 """
 Take the parallax and proper motions of a TGAS star and calculate Jz.
 Append Jz to a csv file.
+Save file as action_data.csv
 """
 
 import numpy as np

### 2017-05-24 chronometer/utils.py
updating parameter names in dict

@@ -28,9 +28,9 @@ def replace_nans_with_inits(data):
         = np.zeros(len(data.feh.values[~np.isfinite(data.feh.values)]))
     data.Av.values[~np.isfinite(data.Av.values)] \
         = np.zeros(len(data.Av.values[~np.isfinite(data.Av.values)]))
-    data.parallax.values[~np.isfinite(data.parallax.values)] \
-        = np.zeros(len(data.parallax.values[~np.isfinite(
-            data.parallax.values)]))
+    data.tgas_parallax.values[~np.isfinite(data.tgas_parallax.values)] \
+        = np.zeros(len(data.tgas_parallax.values[~np.isfinite(
+            data.tgas_parallax.values)]))
     data.tgas_parallax.values[~np.isfinite(data.tgas_parallax.values)] \
         = np.zeros(len(data.tgas_parallax.values[~np.isfinite(
             data.tgas_parallax.values)]))
@@ -64,18 +64,18 @@ def make_param_dict(d, i):
     i: (int)
         The index of the star.
     """
-    param_dict = {"J": (d.j.values[i], d.j_err.values[i]),
-                  "H": (d.h.values[i], d.h_err.values[i]),
-                  "K": (d.k.values[i], d.k_err.values[i]),
+    param_dict = {"J": (d.jmag.values[i], d.jmag_err.values[i]),
+                  "H": (d.hmag.values[i], d.hmag_err.values[i]),
+                  "K": (d.kmag.values[i], d.kmag_err.values[i]),
                   "G": (d.tgas_phot_g_mean_mag.values[i],
                         d.tgas_phot_g_mean_flux_error.values[i]/
                         d.tgas_phot_g_mean_flux.values[i] *
                         d.tgas_phot_g_mean_mag.values[i]),
-                  "Teff": (d.Teff.values[i], d.Teff_err.values[i]),
-                  "logg": (d.logg.values[i], d.logg_err.values[i]),
-                  "feh": (d.feh.values[i], d.feh_err.values[i]),
-                  "parallax": (d.parallax.values[i],
-                               d.parallax_err.values[i])}  # FIXME: add more filters
+                  "Teff": (d.teff.values[i], d.teff_err1.values[i]),
+                  "logg": (d.logg.values[i], d.logg_err1.values[i]),
+                  "feh": (d.feh.values[i], d.feh_err1.values[i]),
+                  "parallax": (d.tgas_parallax.values[i],
[…]

### 2017-05-24 chronometer/chronometer.py
debugging B-V > 0.4

@@ -15,7 +15,7 @@ import h5py
 
 import corner
 import priors
-from models import gc_model, gyro_model
+from models import gc_model, gyro_model, action_age
 import emcee
 
 from utils import replace_nans_with_inits, vk2teff, make_param_dict, \
@@ -76,12 +76,18 @@ def emcee_lnprob(params, *args):
     age_par_inds = par_inds[nglob+N:nglob+2*N]
     g_par_inds = np.concatenate((par_inds[:ngyro],
                                  par_inds[nglob+N:nglob+2*N]))
-    print(params, "params")
-    print(g_par_inds, "g_par_inds")
-    print(params[g_par_inds])
-    assert 0
-    gyro_lnlike = sum(-.5*((period - gyro_model(params[g_par_inds], bv))
-                            /period_errs)**2)
+    m = (bv > .4) * (np.isfinite(period))
+    g_par_inds_mask = np.concatenate((g_par_inds[:3], g_par_inds[3:][m]))
+    print(g_par_inds_mask)
+    gyro_lnlike = sum(-.5*((period[m] - gyro_model(params[g_par_inds_mask],
+                                                     bv[m]))
+                            /period_errs[m])**2)
+    print(bv, "bv")
+    print(period[m], "period")
+    print(-.5*((period[m] - gyro_model(params[g_par_inds], bv[m]))
+               / period_errs[m])**2, "not sum")
+    print(params[g_par_inds], "params")
+    print(gyro_model(params[g_par_inds_mask], bv[m]), "model")
     kin_lnlike = action_age(params[ngyro], params[age_par_inds], Jz, Jz_err)
     p = params*1
     p[nglob:nglob+N] = np.exp(p[nglob:nglob+N])
@@ -89,6 +95,7 @@ def emcee_lnprob(params, *args):
     p[nglob+nglob*N:nglob+4*N] = np.exp(p[nglob+nglob*N:nglob+4*N])
     iso_lnlike = sum([mods[i].lnlike(p[nglob+i::N]) for i in
                       range(len(mods))])
+    print(gyro_lnlike, iso_lnlike, kin_lnlike, lnprior(params, *args))
     return gyro_lnlike + iso_lnlike + kin_lnlike + lnprior(params, *args)
 
 
@@ -342,6 +349,7 @@ if __name__ == "__main__":
     N = len(mods)  # number of stars
     nglob, nind = 4, 5  # number of global pars and number of per-star pars.
     ngyro = 3  # number of gyro parameters.
[…]

### 2017-05-24 chronometer/models.py
debugging B-V > 0.4

@@ -34,6 +34,11 @@ def gyro_model(params, bv):
     a, b, n = params[:3]
     N = len(bv)
     ln_age = params[3:3+N]
+    # p = a*(np.exp(ln_age)*1e3)**n * (bv - .4)**b
+    # mp = np.isfinite(p)
+    # p[~mp] = np.array([None] * len(p[~mp]))  # Using Nones for B-V < .4
+    # return p
+    # mbv = bv > .4
     return a*(np.exp(ln_age)*1e3)**n * (bv - .4)**b
 
 

### 2017-05-24 chronometer/models.py
reverting to previous model

@@ -34,11 +34,6 @@ def gyro_model(params, bv):
     a, b, n = params[:3]
     N = len(bv)
     ln_age = params[3:3+N]
-    # p = a*(np.exp(ln_age)*1e3)**n * (bv - .4)**b
-    # mp = np.isfinite(p)
-    # p[~mp] = np.array([None] * len(p[~mp]))  # Using Nones for B-V < .4
-    # return p
-    # mbv = bv > .4
     return a*(np.exp(ln_age)*1e3)**n * (bv - .4)**b
 
 

### 2017-05-24 chronometer/utils.py
adding mask_column

@@ -9,7 +9,24 @@ from isochrones import StarModel
 from isochrones.mist import MIST_Isochrone
 
 
-def replace_nans_with_inits(data):
+def mask_column(d, col, val):
+    """
+    Turn all the NaNs into the val.
+    params:
+    ------
+    d: (pandas.DataFrame)
+        The data
+    col: (array)
+        The array to mask.
+    val: (float)
+        The value to convert the NaNs into.
+    """
+    m = np.isfinite(col)
+    col[~m] = np.ones(len(col[~m]))*val
+    return col
+
+
+def replace_nans_with_inits(d):
     """
     Turn the data into reasonable initial values.
     params:
@@ -22,22 +39,17 @@ def replace_nans_with_inits(data):
     If no initial Av is provided, initialise with 0.
     """
 
-    data.mass.values[~np.isfinite(data.mass.values)] \
-        = np.ones(len(data.mass.values[~np.isfinite(data.mass.values)]))
-    data.feh.values[~np.isfinite(data.feh.values)] \
-        = np.zeros(len(data.feh.values[~np.isfinite(data.feh.values)]))
-    data.Av.values[~np.isfinite(data.Av.values)] \
-        = np.zeros(len(data.Av.values[~np.isfinite(data.Av.values)]))
-    data.tgas_parallax.values[~np.isfinite(data.tgas_parallax.values)] \
-        = np.zeros(len(data.tgas_parallax.values[~np.isfinite(
-            data.tgas_parallax.values)]))
-    data.tgas_parallax.values[~np.isfinite(data.tgas_parallax.values)] \
-        = np.zeros(len(data.tgas_parallax.values[~np.isfinite(
-            data.tgas_parallax.values)]))
-    data.age.values[~np.isfinite(data.age.values)] \
-        = np.ones(len(data.age.values[~np.isfinite(data.age.values)])) * \
-        np.log(2)
-    return data
+    mass = mask_column(d, d.mass.values, 1.)
+    d["mass"] = mass
+    feh = mask_column(d, d.feh.values, 0.)
+    d["feh"] = feh
+    Av = mask_column(d, d.Av.values, 0.)
+    d["Av"] = Av
+    tgas_parallax = mask_column(d, d.tgas_parallax.values, 0.)
+    d["tgas_parallax"] = tgas_parallax
+    age = mask_column(d, d.age.values, np.log(2))
+    d["age"] = age
+    return d
 
 
 def vk2teff(vk):
[…]

### 2017-05-24 chronometer/priors.py
using any on distance

@@ -29,7 +29,7 @@ def distance_prior(distance, bounds=(0,3000)):
     Distance prior ~ d^2
     """
     min_distance, max_distance = bounds
-    if distance <= min_distance or distance > max_distance:
+    if distance.any() <= min_distance or distance.any() > max_distance:
         return 0
     return 3/max_distance**3 * distance**2
 

### 2017-05-24 chronometer/models.py
log beta

@@ -44,8 +44,9 @@ def action_age(par, t, Jz, Jz_err):
     Vertical action is drawn from a Normal distribution with zero mean and
     dispersion that is a function of time.
     """
-    if par > 0:
-        return np.sum(-.5*(Jz**2/(par*t + Jz_err**2)) - \
-            .5*np.log(2*np.pi*(par*t + Jz_err**2)))
+    beta = np.exp(par)
+    if beta > 0:
+        return np.sum(-.5*(Jz**2/(beta*t + Jz_err**2)) - \
+            .5*np.log(2*np.pi*(beta*t + Jz_err**2)))
     else:
         return -np.inf

### 2017-05-25 chronometer/chronometer.py
running emcee version

@@ -30,6 +30,16 @@ plotpar = {'axes.labelsize': 18,
 plt.rcParams.update(plotpar)
 
 
+def get_n_things(mods, params):
+    """
+    Figure out number of stars, number of global params, etc.
+    """
+    N, nind = len(mods), 5
+    ngyro = 3
+    nglob = len(params) - (N*nind)
+    return N, ngyro, nglob, nind
+
+
 def lnlike(params, *args):
     """
     Probability of age and model parameters given rotation period and colour.
@@ -47,12 +57,19 @@ def lnlike(params, *args):
         The indices of the parameters to vary.
     """
     mods, period, period_errs, bv, bv_errs, par_inds = args
+    N, ngyro, nglob, nind = get_n_things(mods, params)
     gyro_lnlike, iso_lnlike = 0, 0
     if par_inds[0] == 0 and par_inds[1] == 1 and par_inds[2] == 2: # if gyro.
-        r = np.isfinite(period)
-        r_period, r_period_errs, r_bv = period[r], period_errs[r], bv[r]
-        gyro_lnlike = sum(-.5*((r_period - gyro_model(params, r_bv))
-                                /r_period_errs)**2)
+        # r = np.isfinite(period)
+        # r_period, r_period_errs, r_bv = period[r], period_errs[r], bv[r]
+        # gyro_lnlike = sum(-.5*((r_period - gyro_model(params, r_bv))
+                                # /r_period_errs)**2)
+        m = (bv > .4) * (np.isfinite(period))
+        par_inds_mask = np.concatenate((par_inds[:ngyro],
+                                        g_par_inds[ngyro:][m]))
+        gyro_lnlike = sum(-.5*((period[m] - gyro_model(params[par_inds_mask],
+                                                       bv[m]))
+                                /period_errs[m])**2)
 
     # If not gyro but single stars
     else:
@@ -71,23 +88,16 @@ def emcee_lnprob(params, *args):
     The lnprob function used for emcee sampling. Does not break parameters
     into sets.
     """
-    ngyro, nglob, N, mods, period, period_errs, bv, bv_errs, Jz, Jz_err, \
-        par_inds = args
+    N, ngyro, nglob, nind = get_n_things(args[0], params)
[…]

### 2017-05-25 chronometer/chronometer.py
various changes including log prior on beta

@@ -98,7 +98,12 @@ def emcee_lnprob(params, *args):
     gyro_lnlike = sum(-.5*((period[m] - gyro_model(params[g_par_inds_mask],
                                                      bv[m]))
                             /period_errs[m])**2)
-    kin_lnlike = action_age(params[ngyro], params[age_par_inds], Jz, Jz_err)
+
+    # A list of kinematic indices.
+    kin_inds = list(range(nglob+N, nglob+N+nind))
+    kin_inds.insert(0, ngyro)
+    kin_lnlike = action_age(params[kin_inds], Jz, Jz_err)
+
     p = params*1
     p[nglob:nglob+N] = np.exp(p[nglob:nglob+N])
     p[nglob+N:nglob+2*N] = np.log10(1e9*np.exp(p[nglob+N:nglob+2*N]))
@@ -163,7 +168,7 @@ def emcee_lnprior(params, *args):
     mAv = (0 <= params[nglob+4*N:nglob+5*N]) * \
         (params[nglob+4*N:nglob+5*N] < 1)  # Prior on A_v
     m = (-20 < params) * (params < 20)  # Broad bounds on all params.
-    mbeta = 0 < params[ngyro] < 1e20  # Prior on beta
+    mbeta = -20 < params[ngyro] < 20  # Prior on beta
     if sum(m) == len(m) and sum(mAv) == len(mAv) and mbeta:
         return g_prior + age_prior + feh_prior + distance_prior
     else:
@@ -377,9 +382,7 @@ if __name__ == "__main__":
     # Set nsteps and niter.
     nsteps = 1000
     niter = 10
-    N = len(mods)  # number of stars
-    nglob, nind = 4, 5  # number of global pars and number of per-star pars.
-    ngyro = 3  # number of gyro parameters.
+    N, ngyro, nglob, nind = get_n_things(mods, params)
     print(N, "stars")
 
     # Construct parameter indices for the different parameter sets.
@@ -417,12 +420,12 @@ if __name__ == "__main__":
         emcee_args = [mods, d.prot.values,
                       d.prot_err.values, d.bv.values,
                       d.bv_err.values, d.Jz, d.Jz_err, par_inds]
-        nwalkers, nsteps, ndim = 64, 2000, len(params)
+        nwalkers, nsteps, ndim = 64, 10000, len(params)
         p0 = [1e-4*np.random.rand(ndim) + params for i in range(nwalkers)]
[…]

### 2017-05-25 chronometer/emcee_chronometer.py
adding emcee_chronometer

@@ -1,0 +1,154 @@
+"""
+Now use Gibbs sampling to update individual star parameters and global gyro
+parameters.
+"""
+
+import os
+
+import time
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+from isochrones import StarModel
+from isochrones.mist import MIST_Isochrone
+import h5py
+
+import corner
+import priors
+from models import gc_model, gyro_model, action_age
+import emcee
+
+from utils import replace_nans_with_inits, vk2teff, make_param_dict, \
+    parameter_assignment, pars_and_mods, transform_parameters
+
+plotpar = {'axes.labelsize': 18,
+           'font.size': 10,
+           'legend.fontsize': 18,
+           'xtick.labelsize': 18,
+           'ytick.labelsize': 18,
+           'text.usetex': True}
+plt.rcParams.update(plotpar)
+
+
+def get_n_things(mods, params):
+    """
+    Figure out number of stars, number of global params, etc.
+    """
+    N, nind = len(mods), 5
+    ngyro = 3
+    nglob = len(params) - (N*nind)
+    return N, ngyro, nglob, nind
+
+
+def emcee_lnprob(params, *args):
+    """
+    The lnprob function used for emcee sampling. Does not break parameters
+    into sets.
+    """
+    N, ngyro, nglob, nind = get_n_things(args[0], params)
+    mods, period, period_errs, bv, bv_errs, Jz, Jz_err, par_inds = args
+    age_par_inds = par_inds[nglob+N:nglob+2*N]
+    g_par_inds = np.concatenate((par_inds[:ngyro],
+                                 par_inds[nglob+N:nglob+2*N]))
+    m = (bv > .4) * (np.isfinite(period))
+    g_par_inds_mask = np.concatenate((g_par_inds[:3], g_par_inds[3:][m]))
+    gyro_lnlike = sum(-.5*((period[m] - gyro_model(params[g_par_inds_mask],
+                                                     bv[m]))
+                            /period_errs[m])**2)
+
+    # A list of kinematic indices.
+    kin_inds = list(range(nglob+N, nglob+N+nind))
+    kin_inds.insert(0, ngyro)
+    kin_lnlike = action_age(params[kin_inds], Jz, Jz_err)
+
[…]

### 2017-05-25 chronometer/emcee_chronometer.py
adding docstring

@@ -1,6 +1,5 @@
 """
-Now use Gibbs sampling to update individual star parameters and global gyro
-parameters.
+Use emcee to produce posterior samples.
 """
 
 import os

### 2017-05-25 chronometer/chronometer.py
removing emcee from chronometer.py

@@ -83,37 +83,6 @@ def lnlike(params, *args):
     return gyro_lnlike + iso_lnlike
 
 
-def emcee_lnprob(params, *args):
-    """
-    The lnprob function used for emcee sampling. Does not break parameters
-    into sets.
-    """
-    N, ngyro, nglob, nind = get_n_things(args[0], params)
-    mods, period, period_errs, bv, bv_errs, Jz, Jz_err, par_inds = args
-    age_par_inds = par_inds[nglob+N:nglob+2*N]
-    g_par_inds = np.concatenate((par_inds[:ngyro],
-                                 par_inds[nglob+N:nglob+2*N]))
-    m = (bv > .4) * (np.isfinite(period))
-    g_par_inds_mask = np.concatenate((g_par_inds[:3], g_par_inds[3:][m]))
-    gyro_lnlike = sum(-.5*((period[m] - gyro_model(params[g_par_inds_mask],
-                                                     bv[m]))
-                            /period_errs[m])**2)
-
-    # A list of kinematic indices.
-    kin_inds = list(range(nglob+N, nglob+N+nind))
-    kin_inds.insert(0, ngyro)
-    kin_lnlike = action_age(params[kin_inds], Jz, Jz_err)
-
-    p = params*1
-    p[nglob:nglob+N] = np.exp(p[nglob:nglob+N])
-    p[nglob+N:nglob+2*N] = np.log10(1e9*np.exp(p[nglob+N:nglob+2*N]))
-    p[nglob+nglob*N:nglob+4*N] = np.exp(p[nglob+nglob*N:nglob+4*N])
-    iso_lnlike = sum([mods[i].lnlike(p[nglob+i::N]) for i in
-                      range(len(mods))])
-    return gyro_lnlike + iso_lnlike + kin_lnlike + \
-        emcee_lnprior(params, *args)
-
-
 def lnprior(params, *args):
     """
     lnprior on all parameters.
@@ -149,27 +118,6 @@ def lnprior(params, *args):
     m = (-20 < params) * (params < 20)  # Broad bounds on all params.
 
     if sum(m) == len(m) and mAv:
-        return g_prior + age_prior + feh_prior + distance_prior
-    else:
-        return -np.inf
-
-
-def emcee_lnprior(params, *args):
-    """
-    lnprior on all parameters.
-    """
-    N, ngyro, nglob, nind = get_n_things(args[0], params)
-    g_prior = priors.lng_prior(params[:ngyro])
[…]

### 2017-05-26 chronometer/emcee_chronometer.py
fixing bug in prior (why are the bugs always in the prior?)

@@ -44,30 +44,24 @@ def emcee_lnprob(params, *args):
     The lnprob function used for emcee sampling. Does not break parameters
     into sets.
     """
-    N, ngyro, nglob, nind = get_n_things(args[0], params)
-    mods, period, period_errs, bv, bv_errs, Jz, Jz_err, par_inds = args
-    age_par_inds = par_inds[nglob+N:nglob+2*N]
-    g_par_inds = np.concatenate((par_inds[:ngyro],
-                                 par_inds[nglob+N:nglob+2*N]))
-    m = (bv > .4) * (np.isfinite(period))
-    g_par_inds_mask = np.concatenate((g_par_inds[:3], g_par_inds[3:][m]))
+    mods, period, period_errs, bv, bv_errs, Jz, Jz_err, N, ngyro, nglob, \
+        nind, g_par_inds_mask, kin_inds, m = args
+
     gyro_lnlike = sum(-.5*((period[m] - gyro_model(params[g_par_inds_mask],
                                                      bv[m]))
                             /period_errs[m])**2)
 
-    # A list of kinematic indices.
-    kin_inds = list(range(nglob+N, nglob+N+nind))
-    kin_inds.insert(0, ngyro)
     kin_lnlike = action_age(params[kin_inds], Jz, Jz_err)
 
     p = params*1
-    p[nglob:nglob+N] = np.exp(p[nglob:nglob+N])
-    p[nglob+N:nglob+2*N] = np.log10(1e9*np.exp(p[nglob+N:nglob+2*N]))
-    p[nglob+nglob*N:nglob+4*N] = np.exp(p[nglob+nglob*N:nglob+4*N])
+    p[nglob:nglob+N] = np.exp(p[nglob:nglob+N])  # mass
+    p[nglob+N:nglob+2*N] = np.log10(1e9*np.exp(p[nglob+N:nglob+2*N]))  # age
+    p[nglob+3*N:nglob+4*N] = np.exp(p[nglob+3*N:nglob+4*N])  # dist
     iso_lnlike = sum([mods[i].lnlike(p[nglob+i::N]) for i in
                       range(len(mods))])
     return gyro_lnlike + iso_lnlike + kin_lnlike + \
         emcee_lnprior(params, *args)
+    # return gyro_lnlike + emcee_lnprior(params, *args)
 
 
 def emcee_lnprior(params, *args):
@@ -111,20 +105,36 @@ if __name__ == "__main__":
 
     # Construct parameter indices for the different parameter sets.
[…]

### 2017-05-26 chronometer/emcee_chronometer.py
running for longer

@@ -118,7 +118,7 @@ if __name__ == "__main__":
                   d.prot_err.values, d.bv.values, d.bv_err.values, d.Jz,
                   d.Jz_err, N, ngyro, nglob, nind, g_par_inds_mask, kin_inds,
                   m]
-    nwalkers, nsteps, ndim, mult = 64, 100, len(params), 5
+    nwalkers, nsteps, ndim, mult = 64, 1000, len(params), 5
     p0 = [1e-4*np.random.rand(ndim) + params for i in range(nwalkers)]
     sampler = emcee.EnsembleSampler(nwalkers, ndim, emcee_lnprob,
                                     args=emcee_args)
@@ -158,7 +158,9 @@ if __name__ == "__main__":
     fig.savefig(os.path.join(RESULTS_DIR, "emcee_corner"))
 
     # Plot chains
+    print("plotting chains")
     for i in range(ndim):
+        print(i, "of", len(params))
         plt.clf()
         plt.plot(sampler.chain[:, :,  i].T, alpha=.5)
         plt.savefig(os.path.join(RESULTS_DIR, "{}_trace".format(i)))

### 2017-05-26 chronometer/models.py
take ages as params

@@ -37,16 +37,16 @@ def gyro_model(params, bv):
     return a*(np.exp(ln_age)*1e3)**n * (bv - .4)**b
 
 
-def action_age(par, t, Jz, Jz_err):
+def action_age(par, Jz, Jz_err):
     """
     Given a vertical action, calculate an age.
     Vertical action dispersion increases with time.
     Vertical action is drawn from a Normal distribution with zero mean and
     dispersion that is a function of time.
     """
-    beta = np.exp(par)
+    beta, ages = np.exp(par[0]), np.exp(par[1:])
     if beta > 0:
-        return np.sum(-.5*(Jz**2/(beta*t + Jz_err**2)) - \
-            .5*np.log(2*np.pi*(beta*t + Jz_err**2)))
+        return np.sum(-.5*(Jz**2/(beta*ages + Jz_err**2)) - \
+            .5*np.log(2*np.pi*(beta*ages + Jz_err**2)))
     else:
         return -np.inf

### 2017-05-26 chronometer/make_data_file.py
changing period to prot, etc

@@ -29,7 +29,7 @@ if __name__ == "__main__":
     DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
     sun = pd.read_csv(os.path.join(DATA_DIR, "sun.csv"))
     NGC2019 = pd.read_csv(os.path.join(DATA_DIR, "NGC2019.csv"))
-    tgas = pd.read_csv(os.path.join(DATA_DIR, "kic_tgas.csv"))
+    tgas = pd.read_csv(os.path.join(DATA_DIR, "kplr_tgas_periods.csv"))
     tgas = tgas.iloc[:2, :]
 
     # create_and_save_df(sun, NGC2019)
@@ -44,3 +44,6 @@ if __name__ == "__main__":
     # Calculate actions
     action_array = convert_coords_to_actions(df)
     save_pd_file(df, os.path.join(DATA_DIR, "action_data.csv"), action_array)
+
+    print(df["prot"])
+    print(df["tgas_parallax"])

### 2017-05-26 chronometer/test_emcee.py
adding test_emcee.py

@@ -1,0 +1,58 @@
+"""
+Test the emcee functions in chronometer.py
+"""
+
+import os
+import time
+import numpy as np
+import pandas as pd
+from emcee_chronometer import emcee_lnprob, emcee_lnprior, get_n_things
+from utils import pars_and_mods
+
+
+def make_args(mods, params, d):
+    N, ngyro, nglob, nind = get_n_things(mods, params)
+
+    # Construct parameter indices for the different parameter sets.
+    par_inds = np.arange(len(params))  # All
+    g_par_inds = np.concatenate((par_inds[:ngyro],
+                                 par_inds[nglob+N:nglob+2*N]))
+    m = (d.bv.values > .4) * (np.isfinite(d.prot.values))
+    g_par_inds_mask = np.concatenate((g_par_inds[:3], g_par_inds[3:][m]))
+
+    # A list of kinematic indices.
+    kin_inds = list(range(nglob+N, nglob+N+nind))
+    kin_inds.insert(0, ngyro)
+
+    return [mods, d.prot.values, d.prot_err.values, d.bv.values,
+            d.bv_err.values, d.Jz, d.Jz_err, N, ngyro, nglob, nind,
+            g_par_inds_mask, kin_inds, m]
+
+
+def test_emcee_lnprob():
+    DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
+    d = pd.read_csv(os.path.join(DATA_DIR, "action_data.csv"))
+    global_params = np.array([.7725, .601, .5189, np.log(350.)])
+    params, mods = pars_and_mods(d, global_params)
+    emcee_args = make_args(mods, params, d)
+    start = time.time()
+    print(emcee_lnprob(params, *emcee_args))
+    end = time.time()
+    print("time taken = ", end - start, "seconds")
+    start = time.time()
+    print(emcee_lnprob(params, *emcee_args))
+    end = time.time()
+    print("time taken = ", end - start, "seconds")
+
+
+def test_emcee_lnprior():
+    DATA_DIR = "/Users/ruthangus/projects/chronometer/chronometer/data"
+    d = pd.read_csv(os.path.join(DATA_DIR, "action_data.csv"))
+    global_params = np.array([.7725, .601, .5189, np.log(350.)])
+    params, mods = pars_and_mods(d, global_params)
+    emcee_args = make_args(mods, params, d)
+    print(emcee_lnprior(params, *emcee_args))
+
[…]

### 2017-05-26 chronometer/chronometer.py
starting to incorporate into chronometer

@@ -56,20 +56,31 @@ def lnlike(params, *args):
     par_inds: (array) args[-1]
         The indices of the parameters to vary.
     """
-    mods, period, period_errs, bv, bv_errs, par_inds = args
+    mods, period, period_errs, bv, bv_errs, jz, jz_err, par_inds = args
     N, ngyro, nglob, nind = get_n_things(mods, params)
-    gyro_lnlike, iso_lnlike = 0, 0
+    gyro_lnlike, iso_lnlike, kin_lnlike = 0, 0, 0
     if par_inds[0] == 0 and par_inds[1] == 1 and par_inds[2] == 2: # if gyro.
         # r = np.isfinite(period)
         # r_period, r_period_errs, r_bv = period[r], period_errs[r], bv[r]
         # gyro_lnlike = sum(-.5*((r_period - gyro_model(params, r_bv))
                                 # /r_period_errs)**2)
         m = (bv > .4) * (np.isfinite(period))
+        print(m)
         par_inds_mask = np.concatenate((par_inds[:ngyro],
-                                        g_par_inds[ngyro:][m]))
+                                        par_inds[ngyro:][m]))
+        print(par_inds_mask, type(par_inds_mask))
+        print(len(params[par_inds_mask]))
+        print(len(bv[m]), len(period[m]),
+              len(period_errs[m]))
         gyro_lnlike = sum(-.5*((period[m] - gyro_model(params[par_inds_mask],
                                                        bv[m]))
                                 /period_errs[m])**2)
+
+    elif par_inds[0] == 3:  # If kinematics
+        m = np.isfinite(jz)
+        kin_inds_mask = np.concatenate((np.array([par_inds[:1]]),
+                                        par_inds[nglob:][m]))
+        kin_lnlike = action_age(params[kin_inds_mask], Jz, Jz_err)
 
     # If not gyro but single stars
     else:
@@ -80,7 +91,7 @@ def lnlike(params, *args):
         p[1] = np.log10(1e9*np.exp(p[1]))
         p[3] = np.exp(p[3])
         iso_lnlike = ms.lnlike(p)
-    return gyro_lnlike + iso_lnlike
+    return gyro_lnlike + iso_lnlike + kin_lnlike
 
 
 def lnprior(params, *args):
@@ -102,7 +113,7 @@ def lnprior(params, *args):
[…]

### 2017-05-26 chronometer/emcee_chronometer.py
running for longer

@@ -118,7 +118,7 @@ if __name__ == "__main__":
                   d.prot_err.values, d.bv.values, d.bv_err.values, d.Jz,
                   d.Jz_err, N, ngyro, nglob, nind, g_par_inds_mask, kin_inds,
                   m]
-    nwalkers, nsteps, ndim, mult = 64, 1000, len(params), 5
+    nwalkers, nsteps, ndim, mult = 64, 5000, len(params), 5
     p0 = [1e-4*np.random.rand(ndim) + params for i in range(nwalkers)]
     sampler = emcee.EnsembleSampler(nwalkers, ndim, emcee_lnprob,
                                     args=emcee_args)
